\documentclass[10pt,a4paper]{article}
\usepackage[margin=0.62in]{geometry}
\usepackage{array}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{xcolor}
\usepackage{titlesec}

\setlength{\parindent}{0pt}
\setlength{\parskip}{3pt}
\setlist[itemize]{leftmargin=*, itemsep=1pt, topsep=2pt}
\setlist[enumerate]{leftmargin=*, itemsep=1pt, topsep=2pt}
\titleformat{\section}{\large\bfseries\color{blue!45!black}}{}{0em}{}
\titleformat{\subsection}{\normalsize\bfseries\color{black}}{}{0em}{}
\titleformat{\subsubsection}{\normalsize\bfseries\color{black!75}}{}{0em}{}

\newcommand{\monthtag}[1]{\textbf{[#1]}}
\newcommand{\fact}[2]{\item \monthtag{#1} #2}

\begin{document}

\begin{center}
    {\LARGE \textbf{JMI MBA Current Affairs Revision Pack}}\\
    {\large May 2025--April 2026}\\
    \vspace{2mm}
    \textit{Topic-wise exam notes built from monthly current-affairs compilations, with month tags for fast revision.}
\end{center}

\section{How to Use This Pack}
\begin{itemize}
    \item Prioritise the one-line facts, ranks, names, venues, schemes, and rates. These are the most MCQ-friendly items.
    \item Read monthly blocks first, then revise the final topic-wise master lists.
    \item For RBI, Budget, indices, sports tournaments, summits, and awards, memorise exact numbers and pairs.
\end{itemize}

\section{Source Backbone}
\begin{itemize}
    \item Main backbone: The Hindu Review / Adda247 monthly current-affairs PDFs from May 2025 to April 2026.
    \item Cross-check framing: GKToday and Jagran Josh topic structures for exam-style prioritisation.
    \item This is a revision document, not a newspaper archive. Repeated narratives are compressed into factual points.
\end{itemize}

\section{Month-Wise High-Yield Facts}

% BEGIN MAY-AUG 2025
\subsection{May--August 2025: Monthly High-Yield Notes}

\subsubsection{\monthtag{May 2025}}
\begin{itemize}
    \item \textbf{Economy and Banking:} IMF April 2025 WEO projected India as the 4th-largest economy in 2025: India GDP USD 4.187 trillion, Japan USD 4.186 trillion; growth forecast 6.2\% in 2025 and 6.3\% in 2026; top economies: USA, China, Germany, India, Japan.
    \item \textbf{Economy and Banking:} LIC entered Guinness World Records for 5,88,107 life insurance policies sold on 20 January 2025 under ``Mad Million Day''; 4,52,839 agents participated; LIC CEO and MD: Siddhartha Mohanty.
    \item \textbf{Govt Schemes:} National Scheme for ITI Upgradation approved as a Centrally Sponsored Scheme with Rs 60,000 crore outlay; 1,000 govt ITIs, 5 NSTIs, 5 Centres of Excellence, 50,000 trainers, 20 lakh youth in 5 years; ADB and World Bank co-financing.
    \item \textbf{Govt Schemes:} Modified Interest Subvention Scheme continued for FY 2025--26; KCC loans up to Rs 3 lakh at 7\%, 1.5\% subvention to banks, 3\% Prompt Repayment Incentive, effective rate 4\%; covers 7.75 crore farmers.
    \item \textbf{Govt Schemes:} Amrit Bharat Station Scheme: PM inaugurated 103 redeveloped stations across 86 districts in 18 states at over Rs 1,100 crore; overall programme covers 1,300+ stations.
    \item \textbf{National:} President Droupadi Murmu invoked Article 143 seeking SC advisory opinion on Governors'/President's assent powers under Articles 200 and 201; trigger: Tamil Nadu Governor R.N. Ravi delay case and SC use of Article 142.
    \item \textbf{National:} SC mandated minimum 3 years' legal practice for Civil Judge (Junior Division) entry; practice counted from provisional Bar Council enrolment; applies to future recruitment cycles.
    \item \textbf{National Infrastructure:} CCEA approved Ratlam--Nagda 3rd and 4th lines and Wardha--Balharshah 4th line: Rs 3,399 crore, 176 km, 784 villages, 19.74 lakh people, 18.40 MTPA freight, 74 lakh human-days jobs.
    \item \textbf{National Infrastructure:} Badvel--Nellore 4-lane highway in Andhra Pradesh approved: 108.134 km, Rs 3,653.10 crore, DBFOT model; reduces distance by 33.9 km and travel time by about 1 hour.
    \item \textbf{Reports/Static GK:} India's official coastline revised from 7,516.6 km to 11,098.8 km using GIS, satellite altimetry, LIDAR-GPS and drone imaging; measured by Survey of India and National Hydrographic Office; 11 coastal states and 2 UTs.
    \item \textbf{International/Health:} WHO adopted first global pandemic treaty at World Health Assembly, Geneva, on 20 May 2025; vote: 124 in favour, 0 against, 11 abstentions; focus on equity, One Health, pathogen access and benefit sharing.
    \item \textbf{Environment:} Operation Olivia by Indian Coast Guard protected 6.98 lakh Olive Ridley turtles at Rushikulya river mouth, Odisha; annual Nov--May operation started in early 1980s; linked to CMS, CITES and SDG 14.
    \item \textbf{Environment:} 16th Asiatic Lion Census in Gujarat showed 32\% population rise since 2020; Project Lion outlay Rs 2,900 crore approved in March 2025.
    \item \textbf{Science-Tech:} Cabinet approved 6th semiconductor unit under India Semiconductor Mission: HCL-Foxconn JV near Jewar airport, YEIDA region; Rs 3,700 crore investment; 20,000 wafers/month and 36 million units/month.
    \item \textbf{Science-Tech:} Ashwini Vaishnaw inaugurated India's first 3nm semiconductor design hubs by Renesas Electronics India in Noida and Bengaluru; 270+ institutions received EDA tools/learning kits.
    \item \textbf{Science-Tech:} Andhra Pradesh, IBM and TCS announced Quantum Valley Tech Park in Amaravati with India's largest 156-qubit quantum computer, aligned with National Quantum Mission.
    \item \textbf{Defence:} Operation Sindoor: Indian Armed Forces struck 9 terror sites in Pakistan and PoK after Pahalgam terror attack that killed 25 Indians and 1 Nepali; targets linked to LeT, JeM and infiltration hubs; no Pakistani military installations targeted.
    \item \textbf{Defence:} BrahMos production facility inaugurated virtually by Defence Minister Rajnath Singh at UP Defence Industrial Corridor; target production 80--100 BrahMos missiles annually.
    \item \textbf{Awards:} Pulitzer Prize 2025: Public Service -- Kavitha Surana, Lizzie Presser, Cassandra Jaramillo and Stacy Kranitz; Breaking News Reporting -- The Washington Post; Investigative Reporting -- Reuters, ``Fentanyl Express''; NYT won 4 prizes.
    \item \textbf{Awards:} National Florence Nightingale Awards 2025 conferred by President Droupadi Murmu on 15 nursing professionals; award includes certificate, medal and Rs 1,00,000.
    \item \textbf{Sports:} Virat Kohli retired from Test cricket and became fastest cricketer to 13,000 Test runs; Rohit Sharma announced Test retirement; Neeraj Chopra crossed 90 m with 90.23 m at Doha Diamond League.
    \item \textbf{Sports:} R. Vaishali won Asian Individual Chess Championship 2025 in Al-Ain, UAE; Suruchi Singh won gold in 10 m air pistol at ISSF Junior World Cup, Suhl, Germany.
    \item \textbf{Summits/Days:} 10th NITI Aayog Governing Council Meeting held at Bharat Mandapam, New Delhi; theme: ``Viksit Rajya for Viksit Bharat @2047''; attended by 24 states and 7 UTs.
    \item \textbf{Important Days:} World Asthma Day -- 6 May; National Technology Day -- 11 May; International Nurses Day -- 12 May; World Bee Day and World Metrology Day -- 20 May; International Tea Day -- 21 May; World Turtle Day -- 23 May; World No Tobacco Day -- 31 May, theme ``Bright products. Dark intentions.''
    \item \textbf{Static GK:} Lakkundi group of monuments in Karnataka proposed for UNESCO tentative list; Kalyana/Western Chalukya period, 10th--12th century CE; Vesara style; important sites include Kasi Visvesvara, Brahma Jinalaya and Musukina Bavi.
\end{itemize}

\subsubsection{\monthtag{June 2025}}
\begin{itemize}
    \item \textbf{Economy and Banking:} RBI 55th MPC cut repo rate by 50 bps from 6.00\% to 5.50\%; SDF 5.25\%, MSF/Bank Rate 5.75\%; stance shifted from accommodative to neutral; Governor: Sanjay Malhotra.
    \item \textbf{Economy and Banking:} RBI retained FY26 GDP forecast at 6.5\%: Q1 6.5\%, Q2 6.7\%, Q3 6.6\%, Q4 6.3\%; CPI forecast cut to 3.7\% from 4.0\%.
    \item \textbf{Economy and Banking:} MoSPI announced first All-India Household Income Survey in 2026; Technical Expert Group chaired by Surjit Bhalla; NSO to execute; income sources include wages, self-employment, remittances and digital earnings.
    \item \textbf{Govt Schemes:} Digital India completed 11 years; DBT integrated 322+ schemes under 56 ministries, transferred Rs 44 lakh crore and saved Rs 3.48 lakh crore; Aadhaar coverage 140+ crore; DigiLocker 52+ crore users and 852+ crore documents.
    \item \textbf{Govt Schemes:} PMSMA completed 9 years; launched June 2016 by MoHFW; ANC day 9th of every month; 6.19 crore pregnant women examined; E-PMSMA launched Jan 2022 and identified 78.27 lakh high-risk pregnancies by Dec 2024.
    \item \textbf{Govt Schemes:} AMRUT completed 10 years; 2.03 crore tap connections, 1.50 crore sewer connections, 99 lakh LED streetlights; AMRUT 2.0 outlay Rs 2.99 lakh crore with Rs 76,760 crore Central share.
    \item \textbf{National:} Population Census 2027 announced in 2 phases with caste enumeration; reference date 1 March 2027 for most areas and 1 October 2026 for Ladakh and snow-bound areas of J\&K, HP and Uttarakhand; legal basis Census Act, 1948.
    \item \textbf{National Infrastructure:} PM inaugurated Chenab Rail Bridge, completing USBRL; world's highest railway arch bridge, 359 m above Chenab, 467 m arch span; USBRL length 272 km; Anji Bridge is India's first cable-stayed railway bridge.
    \item \textbf{National Infrastructure:} Cabinet approved railway multitracking projects worth Rs 6,405 crore: Koderma--Barkakana 133 km and Ballari--Chikjajur 185 km; total 318 km; 1,408 villages and 28.19 lakh people; 108 lakh human-days jobs.
    \item \textbf{Reports/Indices:} QS World University Rankings 2026: MIT rank 1 for 14th consecutive year; top Indian institution IIT Delhi rank 123, followed by IIT Bombay 129 and IIT Madras 180.
    \item \textbf{Reports/Indices:} Global Peace Index 2025 by Institute for Economics \& Peace: Iceland rank 1, Ireland 2, New Zealand 3; least peaceful Russia rank 163; India rank 115, score 2.229, improved from 116 in 2024.
    \item \textbf{International:} PM Modi received Grand Cross of the Order of Makarios III, Cyprus' highest civilian honour, from President Nikos Christodoulides; first Indian PM visit to Cyprus.
    \item \textbf{International:} India refused to sign SCO Defence Ministers' joint statement in Qingdao because it omitted Pahalgam terror attack; SCO members in 2025 include India, China, Russia, Pakistan, Kazakhstan, Kyrgyzstan, Tajikistan, Uzbekistan, Iran and Belarus; Secretariat Beijing, RATS Tashkent.
    \item \textbf{Environment/Health:} PM launched ``Ek Ped Maa Ke Naam 2.0'' on World Environment Day; target 10 crore trees by 30 September 2025; also flagged off 200 electric buses in Delhi.
    \item \textbf{Environment/Health:} Delhi cloud-seeding pilot: Rs 3.21 crore, 5 aerial trials, silver iodide and salts; support from IIT Kanpur, IMD and national agencies.
    \item \textbf{Environment:} Kerala inaugurated India's first butterfly sanctuary by renaming Aralam Wildlife Sanctuary; 266+ butterfly species in Western Ghats.
    \item \textbf{Science-Tech/Space:} Group Captain Shubhanshu Shukla piloted Axiom-4 to ISS on SpaceX Falcon 9 from Kennedy Space Center on 25 June 2025 at 12:01 pm IST; 2nd Indian in space after Rakesh Sharma and first Indian to live/work on ISS; about 60 experiments, 7 Indian-designed.
    \item \textbf{Science-Tech:} Google opened its first Asia-Pacific Safety Engineering Centre in Hyderabad; focus on online fraud, responsible AI, SynthID watermarking and real-time scam alerts.
    \item \textbf{Defence:} Exercise Tiger Claw: first-ever joint special forces exercise between Indian Air Force Garud Commandos and US special forces in Uttar Pradesh; focus on air-based special operations.
    \item \textbf{Defence:} India-Mongolia Exercise Khaan Quest/Nomadic Elephant references in source: counter-IED, cordon-and-search and civilian evacuation drills; Exercise Shakti with France involved 90-member Indian contingent from Jammu and Kashmir Rifles.
    \item \textbf{Awards:} Bal Sahitya Puraskar 2025 and Sahitya Akademi Yuva Puraskar 2025 announced by Sahitya Akademi under President Madhav Kaushik; English Bal Sahitya winner: Nitin Kushalappa MP for \textit{Dibakar Sinha Indian Myths and Fables Retold}; English Yuva Puraskar: Advait Kolarkar for \textit{Siddhartha -- The Boy Who Became the Buddha}.
    \item \textbf{Awards:} Kerala Sahitya Akademi Awards 2024: Best Novel -- G.R. Indugopan for \textit{Aano}; Poetry -- Anitha Thampi for \textit{Muringa Vazha Kariveppu}; Fellowship -- K.V. Ramakrishnan.
    \item \textbf{Sports:} Royal Challengers Bengaluru won maiden IPL title in IPL 2025 final; Carlos Alcaraz won French Open 2025, his second French Open and fifth Grand Slam; Magnus Carlsen won Norway Chess 2025; D. Gukesh beat Carlsen in a classical game at Norway Chess.
    \item \textbf{Sports:} Neeraj Chopra won Paris Diamond League; Oscar Piastri won Spanish Grand Prix 2025; Lalit Upadhyay retired from international hockey; stop clock introduced in Test cricket.
    \item \textbf{Important Days:} World Milk Day -- 1 June, theme ``Let's Celebrate the Power of Dairy''; World Environment Day -- 5 June, theme on ending plastic pollution; World Food Safety Day -- 7 June; World Blood Donor Day -- 14 June; World Refugee Day -- 20 June; International Olympic Day -- 23 June.
    \item \textbf{Static GK:} International Potato Centre South Asia Regional Centre approved at Singna, Agra; focus on climate-resilient potato and sweetpotato varieties for India and South Asia.
\end{itemize}

\subsubsection{\monthtag{July 2025}}
\begin{itemize}
    \item \textbf{Economy and Banking:} India-UK FTA signed on 24 July 2025 after talks begun in January 2022; eliminates duties on 99\% of Indian exports to UK; India-UK Vision 2035 unveiled; projected annual bilateral trade boost GBP 25.5 billion.
    \item \textbf{Economy and Banking:} FTA tariff changes: Scotch whisky/gin duties to fall from 150\% to 75\% initially and 40\% over 10 years; UK cars duties above 100\% to 10\% under quota; social security exemption for Indian workers in UK up to 3 years, saving approx. Rs 4,000 crore annually.
    \item \textbf{Economy and Banking:} Gujarat became 3rd state after Maharashtra and Uttar Pradesh to cross 1 crore stock market investors; India had nearly 11.5 crore investors as of May 2025.
    \item \textbf{Govt Schemes:} National Sports Policy 2025 approved, replacing 2001 policy; 5 pillars: global excellence, economic development, social development, people's movement, integration with education; vision includes 2036 Olympics.
    \item \textbf{Govt Schemes:} RDI Scheme approved with Rs 1 lakh crore corpus; managed by ANRF; includes concessional loans, equity support and Deep-Tech Fund of Funds for private R\&D in strategic/sunrise sectors.
    \item \textbf{Govt Schemes:} Employment Linked Incentive Scheme approved with Rs 99,446 crore outlay; target 3.5 crore jobs in 2 years; first-time employees get one month's wage up to Rs 15,000; employer incentive Rs 1,000--3,000 per month per additional employee.
    \item \textbf{Govt Schemes:} PM Dhan-Dhaanya Krishi Yojana approved for 2025--31, targeting 100 districts; convergence of 36 schemes across 11 departments; focus on low productivity, low cropping intensity and low credit disbursement.
    \item \textbf{National:} Union Government notified transfer/repatriation of 18 High Court judges on 14 July 2025 after SC Collegium recommendation of 26 May 2025; Article 222 governs transfer after CJI consultation.
    \item \textbf{National:} Bihar approved 35\% reservation in state government jobs only for women domiciled in Bihar; earlier benefit was for all women.
    \item \textbf{National:} Bihar announced free electricity up to 125 units/month from 1 August 2025 for 16.7 million families; under Kutir Jyoti, solar panels for poor households with 10,000 MW target in 3 years.
    \item \textbf{Reports/Indices:} UN SDG Index 2025: Finland rank 1 score 87.0; Sweden 2, Denmark 3; India rank 99, score 67.0, first time in top 100; South Sudan rank 167, score 41.6.
    \item \textbf{Reports/Indices:} Swachh Survekshan 2024--25: Ahmedabad cleanest city above 10 lakh population; 4,589 cities covered; Super Swachh League included Indore, Surat and Navi Mumbai among >10 lakh cities.
    \item \textbf{International:} PM Modi received Officer of the Order of the Star of Ghana from President John Dramani Mahama; first Indian PM visit to Ghana in over 30 years; India to offer Bharat UPI and Jan Aushadhi Kendras support.
    \item \textbf{International:} PM Modi received Order of the Republic of Trinidad and Tobago on 4 July 2025; India and Trinidad \& Tobago signed 6 MoUs: pharma standards, quick impact projects, culture, sports, diplomatic training and document/research sharing; OCI access extended to 6th-generation Indian origin people.
    \item \textbf{International:} PM Modi received Namibia's Order of the Most Ancient Welwitschia Mirabilis from President Netumbo Nandi-Ndaitwah; first Indian leader to receive it; also received Brazil's Grand Collar of the National Order of the Southern Cross.
    \item \textbf{International:} PM Modi received ``Key to the City of Buenos Aires'' from Jorge Macri; first Indian PM visit to Argentina in 57 years; talks with President Javier Milei on lithium, shale gas, defence, energy, agriculture, mining and trade.
    \item \textbf{Environment/Health:} Tamil Nadu declared Dhanushkodi Greater Flamingo Sanctuary over 524.7 hectares in Gulf of Mannar Biosphere Reserve; 10,700+ birds from 128 species recorded.
    \item \textbf{Environment/Health:} Tamil Nadu approved Centre of Excellence for Hornbill Conservation at Anamalai Tiger Reserve with Rs 1 crore; targets 4 hornbill species.
    \item \textbf{Science-Tech:} BITS Pilani to set up India's first ``AI Plus Campus'' in Amaravati, Andhra Pradesh; Rs 1,000 crore project; admissions from 2027; focus AI, Data Science and Robotics.
    \item \textbf{Science-Tech:} Himachal Pradesh launched India's first Aadhaar-based FaceAuth system for ration distribution under PDS; uses smartphone camera instead of OTP/fingerprint.
    \item \textbf{Defence:} India-UK Vision 2035 includes Defence Industrial Roadmap, Technology Security Initiative, climate action and education collaboration.
    \item \textbf{Defence:} Mumbai's 328-metre Carnac Bridge, built in 1868, was rebuilt and renamed Sindoor Bridge to honour Operation Sindoor.
    \item \textbf{Sports:} Divya Deshmukh won FIDE Women's Chess World Cup 2025 in Batumi, Georgia, defeating Koneru Humpy in tiebreaks; became India's 4th woman Grandmaster after Koneru Humpy, Harika Dronavalli and R. Vaishali.
    \item \textbf{Sports:} Jannik Sinner won first Wimbledon title on 13 July 2025; Iga Swiatek won first Wimbledon title on 12 July 2025; D. Gukesh won rapid title at Grand Chess Tour, Zagreb; Lando Norris won British Grand Prix at Silverstone.
    \item \textbf{Awards/Appointments:} Sanjog Gupta appointed ICC CEO; President Christine Carla Kangaloo of Trinidad \& Tobago earlier received Pravasi Bharatiya Samman Award.
    \item \textbf{Summits/Important Days:} National Turmeric Board HQ inaugurated in Nizamabad, Telangana, by Amit Shah; Rs 200 crore funding; export goal USD 1 billion by 2030.
    \item \textbf{Important Days:} World Population Day -- 11 July, theme ``Empowering young people to create the families they want in a fair and hopeful world''; World Youth Skills Day -- 15 July, theme ``Youth Empowerment Through Artificial Intelligence and Digital Skills''; World Chess Day -- 20 July; World Mangrove Day -- 26 July; World Hepatitis Day -- 28 July, theme ``Hepatitis: Let's Break It Down''.
    \item \textbf{Static GK:} Maratha Military Landscapes of India added to UNESCO World Heritage List at 47th WHC session in Paris; 12 forts: Salher, Shivneri, Lohgad, Khanderi, Raigad, Rajgad, Pratapgad, Suvarnadurg, Panhala, Vijay Durg, Sindhudurg in Maharashtra and Gingee in Tamil Nadu.
\end{itemize}

\subsubsection{\monthtag{August 2025}}
\begin{itemize}
    \item \textbf{Economy and Banking:} RBI 56th MPC kept repo rate unchanged at 5.50\%; SDF 5.25\%, MSF/Bank Rate 5.75\%; FY26 GDP forecast 6.5\%, CPI forecast 3.1\%; June CPI 2.1\%, 77-month low; food inflation -0.2\%.
    \item \textbf{Economy and Banking:} Revised Income Tax Bill, 2025 introduced after withdrawal of earlier draft; based on about 285 Select Committee recommendations; 536 sections and 16 schedules; introduces single ``Tax Year'' and digital-first administration; no immediate slab changes.
    \item \textbf{Economy and Banking:} Key tax changes proposed: refund claims even for late returns, Section 80M inter-corporate dividend deduction, NIL-TDS certificates, removal of deemed rent on vacant properties, 30\% house-property deduction after municipal taxes, home-loan interest deduction for let-out properties.
    \item \textbf{Economy and Banking:} Barclays Private Clients Hurun India Most Valuable Family Businesses 2025: Ambani family rank 1, Reliance Industries, Rs 28.23 lakh crore; Birla rank 2, Rs 6.48 lakh crore; Jindal rank 3, Rs 5.71 lakh crore.
    \item \textbf{Economy and Banking:} RBI FREE-AI framework report chaired by Dr Pushpak Bhattacharyya; 7 Sutras: Trust, People First, Innovation over Restraint, Fairness \& Equity, Accountability, Understandable by Design, Safety/Resilience/Sustainability; only 20.8\% supervised entities use/develop AI.
    \item \textbf{Govt Schemes:} Online Gaming Bill, 2025 passed by Lok Sabha; promotes e-sports and social games, bans online money games; violation can bring up to 3 years imprisonment and Rs 1 crore fine; repeat offences 3--5 years and up to Rs 2 crore fine.
    \item \textbf{Govt Schemes:} Cabinet extended/restructured PM SVANidhi till 31 March 2030 with Rs 7,332 crore outlay; 1.15 crore vendors including 50 lakh new; loans Rs 15,000, Rs 25,000, Rs 50,000; UPI-linked RuPay credit cards and cashback up to Rs 1,600.
    \item \textbf{Govt Schemes:} Cabinet approved MERITE Scheme, Rs 4,200 crore outlay for 2025--30, supported by Rs 2,100 crore World Bank loan; 275 institutions and 7.5 lakh students.
    \item \textbf{Govt Schemes:} PMUY subsidy approved: Rs 300 per 14.2 kg LPG cylinder, up to 9 refills in FY 2025--26; outlay Rs 12,000 crore; 10.33 crore households.
    \item \textbf{National:} PM's 79th Independence Day address lasted 103 minutes, the longest from Red Fort; announced Mission Sudarshan Chakra security shield by 2035, GST reform by Diwali 2025 and PM Viksit Bharat Rozgar Yojana of Rs 1 lakh crore for 3 crore beneficiaries.
    \item \textbf{National:} PM highlighted National Critical Minerals Mission with exploration of 1,200 sites, National Deepwater Exploration Mission, 50\% clean energy target achieved in 2025, and nuclear capacity target 10-fold by 2047.
    \item \textbf{Reports/Indices:} Numbeo Safety Index mid-2025: India rank 67 globally, score 55.8; safest Indian city Mangalore score 74.2, global rank 49; global top city Abu Dhabi score 88.8 for 9th consecutive year.
    \item \textbf{Reports/Indices:} GeM crossed Rs 15 lakh crore GMV; EPFO recorded highest-ever net addition of 21.89 lakh members in June 2025; India ranked 2nd globally in fish production.
    \item \textbf{International:} PM Modi visited Japan on 29--30 August for 15th India-Japan Annual Summit with PM Ishiba Shigeru; adopted India-Japan Joint Vision for Next Decade to 2035 covering 8 sectors.
    \item \textbf{International:} India-Japan outcomes: Joint Declaration on Security Cooperation, Digital Partnership 2.0, AI Initiative, Economic Security Initiative, Critical Minerals MoU, Clean Hydrogen and Ammonia declaration, Chandrayaan-5/LUPEX lunar polar mission arrangement.
    \item \textbf{International:} India-Japan set target of Yen 10 trillion Japanese private investment over next decade; HR exchange plan for 5,00,000 personnel including 50,000 skilled Indians in 5 years; Japan to introduce Shinkansen E10 trains for Mumbai-Ahmedabad HSR.
    \item \textbf{Environment/Health:} Cabinet approved 700 MW Tato-II Hydro Electric Project in Shi Yomi, Arunachal Pradesh: Rs 8,146.21 crore, 72 months, annual generation 2,738 MU, 12\% free power to state.
    \item \textbf{Environment/Health:} Kerala declared India's first fully digitally literate state under Digi Kerala, training 21.87 lakh citizens; Delhi Assembly became first in India to run entirely on solar power with 500 kW rooftop plant.
    \item \textbf{Science-Tech:} Semiconductor projects worth Rs 4,600 crore approved in Odisha, Punjab and Andhra Pradesh: India's first commercial compound fab by SiCSem in Odisha, 3D advanced packaging unit in Odisha, ASIP facility in Andhra Pradesh and CDIL expansion in Punjab; 2,034 direct jobs.
    \item \textbf{Science-Tech:} CG Power inaugurated OSAT pilot line facility at Sanand, Gujarat under India Semiconductor Mission; Rs 7,600 crore investment and 5,000+ jobs expected.
    \item \textbf{Science-Tech:} India achieved 100 GW solar PV module manufacturing capacity under ALMM; Deendayal Port Authority commissioned India's first indigenous 1 MW green hydrogen plant at Kandla with 10 MW expansion plan.
    \item \textbf{Science-Tech:} India Post launched IT 2.0 -- Advanced Postal Technology with DIGIPIN, QR payments and OTP delivery; SabhaSaar AI tool launched for Gram Sabha minutes using BHASHINI.
    \item \textbf{Defence:} Operation Sindoor gallantry awards announced before Independence Day: Sarvottam Yudh Seva Medal to Air Marshal Narnadeshwar Tiwari, Air Marshal Jeetendra Mishra, Air Marshal Awadhesh Bharti, Vice Admiral S.J. Singh and two senior Army officers; UYSM to Vice Admiral Tarun Sobti.
    \item \textbf{Defence:} Operation Sindoor involved Army destruction of 9 terrorist bases linked to JeM, LeT and Hizbul Mujahideen; IAF precision strikes on 4 targets inside Pakistan and 5 in PoK; Navy operational support from Indian soil.
    \item \textbf{Defence:} National Sports Governance Bill and National Anti-Doping (Amendment) Bill passed by Parliament; NCERT launched Operation Sindoor modules for schools.
    \item \textbf{Sports:} Cabinet approved India's bid to host Commonwealth Games 2030 in Ahmedabad; if accepted, Gujarat to receive grant-in-aid; expected participation from 72 countries; Narendra Modi Stadium hosted 2023 ICC Cricket World Cup Final.
    \item \textbf{Sports:} Mumbai hosted 18th International Olympiad on Astronomy and Astrophysics with about 300 students from 64 countries and 12 first-time participant countries.
    \item \textbf{Awards:} Educate Girls/Foundation to Educate Girls Globally became first Indian organisation to win Ramon Magsaysay Award 2025; founded in 2007 by Safeena Husain; 2 million+ girls enrolled/retained, 30,000+ villages, retention above 90\%.
    \item \textbf{Awards:} Other Ramon Magsaysay 2025 awardees named in source: Shaahina Ali of Maldives for environmental conservation and Flaviano Antonio L. Villanueva of Philippines for marginalised communities; award founded in 1957 and presented in Manila on 7 November.
    \item \textbf{Summits/Important Days:} India-Japan Year of Science, Technology and Innovation Exchange 2025 marked 40 years of S\&T cooperation; 75th diplomatic anniversary due in 2027.
    \item \textbf{Important Days:} World Breastfeeding Week theme: ``Prioritise Breastfeeding: Create Sustainable Support Systems''; World Wide Web Day theme: ``Empowering the Future: Building an Inclusive, Safe, and Open Web''; Maharashtra to observe 7 August, birth anniversary of Dr M.S. Swaminathan, as Sustainable Agriculture Day.
    \item \textbf{Static GK:} Sawalkote Hydropower Project in J\&K: 2,185 MW run-of-river plant on Chenab, revived by NHPC and JKSPDC at Rs 22,704 crore; target annual generation 7,000 million units.
    \item \textbf{Static GK:} Cabinet cleared 4 railway projects worth Rs 12,328 crore adding 565 km tracks, including 145 km new line in Kutch connecting Rann of Kutch, Dholavira, Koteshwar and Narayan Sarovar; expected freight boost 68 MTPA.
\end{itemize}
% END MAY-AUG 2025

% BEGIN SEP-DEC 2025
\subsection{September--December 2025: Monthly High-Yield Notes}

\subsubsection{\monthtag{September 2025}}
\begin{itemize}
    \item \textbf{Economy and Banking:} GST reforms simplified slabs mainly to 5\% and 18\%; essentials/education items moved to 0--5\%, individual health and life insurance from 18\% to nil, small cars/two-wheelers up to 350 cc from 28\% to 18\%.
    \item \textbf{Economy and Banking:} GST revenue in August 2025 reached Rs 1.86 trillion, up 6.5\% YoY; retail inflation in August was 2.07\%, below RBI target; WPI inflation turned positive at 0.52\%.
    \item \textbf{Economy and Banking:} India's CAD narrowed to USD 2.4 billion, 0.2\% of GDP, in Q1 FY26; services receipts and 18\% remittance rise helped; trade deficit in August was USD 26.49 billion.
    \item \textbf{Economy and Banking:} Morgan Stanley raised India's FY26 GDP forecast to 6.7\%; S\&P retained FY26 at 6.5\%; Fitch raised FY25 forecast to 6.9\%; services PMI touched 62.9 in August, a 15-year high.
    \item \textbf{Banking/Finance:} RBI permitted working-capital loans using gold as raw material from 1 October 2025; mandated 2-factor authentication for all digital payments from 1 April 2026.
    \item \textbf{Banking/Finance:} NPCI raised UPI limits to Rs 5 lakh per transaction and Rs 10 lakh daily for select verified merchants; UPI went live in Qatar, the 8th country to adopt it for Indian travellers.
    \item \textbf{Govt Schemes:} Cabinet approved Rs 1,500 crore Critical Mineral Recycling Scheme for FY 2025--26 to 2030--31; target 270 KT annual recycling capacity, 40 KT critical minerals, Rs 8,000 crore investment and 70,000 jobs.
    \item \textbf{Govt Schemes:} Bihar's Mukhyamantri Mahila Rojgar Yojana gives Rs 10,000 DBT support to 75 lakh women for self-employment; Tamil Nadu launched Anbu Karangal for orphaned children.
    \item \textbf{National Infrastructure:} CCEA cleared Rs 4,447.38 crore Mokama--Munger 82.4 km greenfield corridor, Rs 3,822.31 crore Sahebganj--Areraj--Bettiah NH-139W section, and Rs 2,192 crore Bakhtiyarpur--Rajgir--Tilaiya rail doubling.
    \item \textbf{National:} C.P. Radhakrishnan elected India's 15th Vice President on 9 September; secured 452 first-preference votes against Justice B. Sudershan Reddy's 300; turnout 98.20\%.
    \item \textbf{Reports/Indices:} Global Innovation Index 2025 by WIPO: India rank 38, up from 48 in 2020; Switzerland rank 1, followed by Sweden, USA, Korea and Singapore.
    \item \textbf{Reports/Indices:} NIRF Rankings 2025: overall top 3 were IIT Madras, IISc Bengaluru and IIT Bombay; Management top was IIM Ahmedabad; Law top was NLSIU Bengaluru; Pharmacy top was Jamia Hamdard.
    \item \textbf{Reports/Indices:} Bloomberg Billionaires Index, 11 September: Elon Musk USD 384 billion rank 1; Larry Ellison USD 383 billion rank 2 after USD 89 billion one-day gain; Mukesh Ambani rank 18, Gautam Adani rank 21.
    \item \textbf{International:} 25th SCO Summit held in Tianjin; PM Modi stressed Security, Connectivity and Opportunity; Tianjin Declaration condemned the 22 April Pahalgam terror attack.
    \item \textbf{International:} Nepal's Gen-Z protests after a social-media ban led to KP Sharma Oli's resignation; Sushila Karki was sworn in on 12 September as Nepal's first woman Prime Minister/interim leader.
    \item \textbf{International:} India-Bhutan approved first rail links, Kokrajhar--Gelephu and Banarhat--Samtse, worth Rs 4,033 crore; India-EFTA TEPA to operate from 1 October 2025.
    \item \textbf{Environment/Health:} Cold Desert Biosphere Reserve, Himachal Pradesh, joined UNESCO World Network as India's 13th biosphere reserve; Bihar added 2 Ramsar sites, taking India's total to 93.
    \item \textbf{Environment/Health:} India achieved 250 GW non-fossil capacity against 500 GW target by 2030; MNRE released India's first National Geothermal Policy with 10 GW potential.
    \item \textbf{Science-Tech:} Vikram-3201, India's first indigenous 32-bit microprocessor for space use, was unveiled; Adi Vaani AI Translator launched for Santali, Bhili, Mundari and Gondi.
    \item \textbf{Science-Tech:} Semicon India 2025 in New Delhi showcased 10 projects worth Rs 1.5 lakh crore with 48+ countries; India nominated IIT Madras as UN AI Centre of Excellence under ODET.
    \item \textbf{Defence:} Yudh Abhyas 2025, 21st India-US edition, held at Fort Wainwright, Alaska from 1--14 September; Yudh Kaushal 3.0 held in Kameng, Arunachal Pradesh with drones, AI tactics and indigenous weapons.
    \item \textbf{Defence:} Operation Rahat by Army evacuated 6,000+ flood-hit people and gave medical aid to 13,000+; indigenous encrypted 5G SAMBHAV system replaced WhatsApp in Operation Sindoor.
    \item \textbf{Sports:} India won 9th Asia Cup title by beating Pakistan by 5 wickets at Dubai International Stadium on 28 September; Tilak Varma 69 not out was Man of the Match; Kuldeep Yadav MVP; Abhishek Sharma Player of Tournament.
    \item \textbf{Awards/Appointments:} 71st National Film Awards at Vigyan Bhawan: Dadasaheb Phalke Award to Mohanlal; Best Feature Film \textit{12th Fail}; Best Actors Shah Rukh Khan and Vikrant Massey; Best Actress Rani Mukerji.
    \item \textbf{Awards/Appointments:} 2025 Ramon Magsaysay Award: Educate Girls became first Indian organisation to win; other awardees included Shaahina Ali and Flaviano Antonio L. Villanueva.
    \item \textbf{Summits/Days/Static GK:} World Food India 2025 held at Bharat Mandapam, New Delhi, 25--28 September; 1700+ exhibitors, 500+ global buyers, 100+ countries; Make in India completed 11 years since 2014.
\end{itemize}

\subsubsection{\monthtag{October 2025}}
\begin{itemize}
    \item \textbf{Economy and Banking:} RBI MPC kept repo rate unchanged at 5.50\%, stance neutral; SDF 5.25\%, MSF/Bank Rate 5.75\%, CRR 3\%; FY26 GDP forecast raised to 6.8\%, CPI forecast cut to 2.6\%.
    \item \textbf{Economy and Banking:} FY26 GDP quarterly projections: Q1 6.5\%, Q2 7.0\%, Q3 6.4\%, Q4 6.2\%; CPI projections: Q2 2.1\%, Q3 3.1\%, Q4 4.4\%.
    \item \textbf{Economy and Banking:} GST collections for September 2025 hit Rs 1.89 lakh crore, up 9.1\% YoY; retail inflation fell to 1.54\%, an 8-year low, with food inflation at -2.28\%.
    \item \textbf{Banking/Finance:} Banking Laws (Amendment) Act 2025 nomination rules effective 1 November allow up to four nominees for deposits; RBI set up Payments Regulatory Board chaired by Governor Sanjay Malhotra.
    \item \textbf{Banking/Finance:} UPI launched in Qatar at Lulu supermarkets through NIPL-QNB-NETSTARS; NPCI launched biometric UPI authentication and AI-powered UPI Help; PayPal integrated UPI as first global partner in PayPal World.
    \item \textbf{Govt Schemes:} Cabinet approved Mission for Aatmanirbharta in Pulses, outlay Rs 11,440 crore for 2025--31; target 350 lakh tonnes production and 310 lakh hectares by 2030--31.
    \item \textbf{Govt Schemes:} PM SETU Scheme: Rs 60,000 crore to modernise 1,000 ITIs and benefit 2 million youth over 5 years; Dhan-Dhaanya Krishi Yojana: Rs 24,000 crore for agricultural productivity.
    \item \textbf{Govt Schemes:} Cabinet approved BRCP Phase-III, 2025--26 to 2030--31, cost Rs 1,500 crore, DBT Rs 1,000 crore and Wellcome Trust Rs 500 crore; target 2,000+ researchers.
    \item \textbf{National Infrastructure:} CCEA approved NH-715 Kalibor--Numaligarh four-laning in Assam, 85.675 km, Rs 6,957 crore, including 34.45 km elevated wildlife corridor through Kaziranga landscape.
    \item \textbf{National Infrastructure:} Cabinet approved 57 new Kendriya Vidyalayas at Rs 5,862.55 crore over 9 years and four railway multitracking projects worth Rs 24,634 crore covering 894 km.
    \item \textbf{Reports/Indices:} Global Firepower Index 2025: USA rank 1, Russia 2, China 3, India 4; India's PowerIndex score 0.1184.
    \item \textbf{Reports/Indices:} World's most valuable companies, October 2025: NVIDIA USD 4.894 trillion rank 1, Microsoft USD 4.029 trillion rank 2, Apple USD 3.992 trillion rank 3; Reliance rank 69 globally, USD 230.69 billion.
    \item \textbf{International:} Timor-Leste became ASEAN's 11th member at Kuala Lumpur Summit on 26 October; East Timor's membership process took 14 years.
    \item \textbf{International:} UK PM Keir Starmer's India visit produced 12 outcomes including Joint AI Centre, Connectivity \& Innovation Centre and UK university campuses in Bengaluru and GIFT City.
    \item \textbf{International:} Sanae Takaichi became Japan's first female Prime Minister; Catherine Connolly elected Ireland's 10th President with 63\% votes; Rodrigo Paz won Bolivia presidency with 54.5\%.
    \item \textbf{Environment/Health:} Project Dolphin Phase II, Project Sloth Bear and Project Gharial launched during Wildlife Week 2025; India also launched National Red List Roadmap, Vision 2025--2030.
    \item \textbf{Environment/Health:} Palk Bay, Tamil Nadu, recognised as India's first dugong reserve by IUCN; India ranked 9th in total forest area and 3rd in annual forest gain in FAO GFRA 2025.
    \item \textbf{Science-Tech:} Nobel Medicine 2025: Mary E. Brunkow, Fred Ramsdell and Shimon Sakaguchi for peripheral immune tolerance; Physics: John Clarke, Michel H. Devoret and John M. Martinis for macroscopic quantum tunnelling.
    \item \textbf{Science-Tech:} Nobel Chemistry 2025: Susumu Kitagawa, Richard Robson and Omar M. Yaghi for metal-organic frameworks; Literature: Laszlo Krasznahorkai; Peace: Maria Corina Machado; Economics: Joel Mokyr, Philippe Aghion and Peter Howitt.
    \item \textbf{Science-Tech:} India-UK launched GBP 24 million Connectivity \& Innovation Centre for AI telecom, 6G and cybersecurity; Global Fintech Fest 2025 held in Mumbai, theme ``Empowering Finance for a Better World Powered by AI''.
    \item \textbf{Defence:} India-US signed a 10-year defence framework in Kuala Lumpur on 31 October covering joint exercises, technology sharing and Indo-Pacific maritime security.
    \item \textbf{Defence:} President Droupadi Murmu became first Indian President to fly in a Rafale fighter jet, completing a 30-minute sortie from Ambala Air Force Station.
    \item \textbf{Sports:} Abhishek Sharma and Smriti Mandhana named ICC Players of the Month for September/October source cycle; Abhishek scored 314 runs in Asia Cup at strike rate 200; Mandhana hit India's fastest women's ODI century in 50 balls.
    \item \textbf{Awards/Appointments:} 70th Filmfare Awards 2025 at EKA Arena, Ahmedabad: \textit{Laapataa Ladies} won 13 awards incl. Best Film and Best Director Kiran Rao; Alia Bhatt Best Actress for \textit{Jigra}.
    \item \textbf{Awards/Appointments:} Justice Surya Kant named 53rd CJI from 23 November 2025 with tenure till February 2027; Neeraj Chopra conferred honorary Lieutenant Colonel rank in Territorial Army.
    \item \textbf{Summits/Days/Static GK:} 150 years of \textit{Vande Mataram} noted; composed by Bankim Chandra Chattopadhyay and first sung by Rabindranath Tagore in 1896; Janajatiya Gaurav Diwas falls on 15 November for Birsa Munda.
\end{itemize}

\subsubsection{\monthtag{November 2025}}
\begin{itemize}
    \item \textbf{Economy and Banking:} UPI transactions in October 2025 reached Rs 27.28 lakh crore across 20.7 billion transactions; UPI operational in 7 countries as per source.
    \item \textbf{Economy and Banking:} India surpassed Japan to become world's 3rd-largest equity market, crossing USD 5 trillion; merchandise plus services exports in Apr--Oct 2025 touched USD 491.8 billion, up 4.84\%.
    \item \textbf{Economy and Banking:} Public sector banks posted Rs 49,456 crore profit in Q2 FY26, up 9\% YoY; SBI contributed Rs 20,160 crore and crossed USD 100 billion market cap plus Rs 100 trillion business milestone.
    \item \textbf{Economy and Banking:} India signed first structured LPG import agreement with US suppliers for 2.2 MTPA from 2026, covering about 10\% of demand.
    \item \textbf{Govt Schemes:} Four labour codes came into effect on 21 November 2025, consolidating 29 laws: Code on Wages 2019, Industrial Relations Code 2020, Social Security Code 2020, OSHWC Code 2020.
    \item \textbf{Govt Schemes:} Labour-code facts: 8-hour day/48-hour week, overtime at 2x wage, fixed-term gratuity after 1 year, layoff approval threshold raised to 300 workers, gig/platform workers covered.
    \item \textbf{Govt Schemes:} Cabinet approved Export Promotion Mission with Rs 25,060 crore outlay; two components: NIRYAT PROTSAHAN for trade finance and NIRYAT DISHA for compliance, branding, logistics and market access.
    \item \textbf{Govt Schemes:} Cabinet approved 100\% credit guarantee via NCGTC for additional lending up to Rs 20,000 crore under Credit Guarantee Scheme for Exporters.
    \item \textbf{Govt Schemes:} Cabinet approved Rs 7,280 crore scheme for Sintered Rare Earth Permanent Magnets, creating India's first integrated 6,000 MTPA REPM ecosystem.
    \item \textbf{National:} Nitish Kumar sworn in as Bihar CM for record 10th term after 2025 Assembly elections; Justice Surya Kant took oath as 53rd CJI, tenure till February 2027.
    \item \textbf{National:} MeitY notified DPDP Rules 2025, operationalising Digital Personal Data Protection Act 2023; National Cyber Shield launched as AI-enabled cyber-defence network integrating 50+ agencies.
    \item \textbf{Reports/Indices:} India Skills Report 2026 by ETS, CII, AICTE, AIU and Taggd: employability rose to 56.35\% from 54.81\%; women 54\% outperformed men 51.5\%; 22--25 age group at 75.7\%.
    \item \textbf{Reports/Indices:} QS World University Sustainability Rankings 2026: Lund University, Sweden rank 1; University of Toronto rank 2; UCL rank 3; India had 103 universities and 26 debutants.
    \item \textbf{Reports/Indices:} World Justice Project Rule of Law Index 2025: India rank 86 of 143; Denmark topped; G20 Inequality Report noted India's top 1\% own 27\% of national wealth.
    \item \textbf{International:} President Droupadi Murmu made first-ever state visit by an Indian President to Botswana, 11--13 November; Botswana became 18th country to recognise Indian Pharmacopoeia.
    \item \textbf{International:} Botswana donated eight cheetahs for Project Cheetah; to be sent to Kuno National Park, Madhya Pradesh; cheetahs became extinct in India in 1952.
    \item \textbf{International:} 20th G20 Summit held in Johannesburg, 22--23 November; theme ``Solidarity, Equality, Sustainability''; US did not participate; 122-point Leaders' Declaration adopted.
    \item \textbf{International/Summits:} PM Modi proposed six G20 ideas: Traditional Knowledge Repository, Africa Skills Multiplier, Global Health Response Team, Open Satellite Data Partnership, Critical Minerals Circularity Initiative, countering drug-terror nexus.
    \item \textbf{Environment/Health:} COP30 held in Belem, Brazil, 10--21 November; Belem Package adopted with 29 decisions; India joined TFFF as observer; Global Big Cats Summit 2026 announced for New Delhi.
    \item \textbf{Environment/Health:} ARISE programme launched by Climate Investment Fund for LDCs and SIDS; Germany USD 63.25 million and Spain USD 36.8 million, about USD 100 million total.
    \item \textbf{Science-Tech:} India launched BIRSA 101, first indigenous CRISPR-based gene therapy for sickle cell disease; first MWh-scale Vanadium Flow Battery unveiled at NTPC NETRA.
    \item \textbf{Science-Tech:} National Quantum Mission facilities set up at IIT Bombay, IISc Bengaluru, IIT Delhi and IIT Kanpur; India's first Liquid Helium Facility inaugurated at IIT Bombay.
    \item \textbf{Defence/Security:} Cabinet condemned 10 November terrorist car explosion near Red Fort, Delhi; government reiterated zero tolerance to terrorism and ordered urgent investigation.
    \item \textbf{Defence/Security:} US approved supply to India of 100 FGM-148 Javelin ATGMs worth USD 45.7 million and 216 Excalibur M982A1 artillery rounds worth USD 47.1 million.
    \item \textbf{Sports:} ISSF World Championship Rifle/Pistol 2025 held in Cairo, 6--18 November; India finished 3rd with 13 medals: 3 gold, 6 silver, 4 bronze; China topped with 21 medals.
    \item \textbf{Sports:} Lakshya Sen won men's singles at Australian Open 2025 Badminton, Sydney, defeating Yushi Tanaka 21-15, 21-11; event was BWF Super 500 with USD 475,000 prize money.
    \item \textbf{Awards/Appointments:} 53rd International Emmy Awards: Best Drama Series \textit{Rivals}; Comedy \textit{Ludwig}; Actress Anna Maxwell Martin; Actor Oriol Pla; Directorate Award Joao Roberto Marinho; Founder's Award Dana Walden.
    \item \textbf{Awards/Appointments:} Shiv Nadar family topped India's philanthropists list with Rs 2,708 crore FY25 donation; Mukesh Ambani family second Rs 626 crore; Bajaj family third Rs 446 crore.
    \item \textbf{Static GK:} G20 established 1999; first Leaders' Summit 2008; members: 19 countries plus EU and African Union; represents about 85\% global GDP, 75\% world trade and two-thirds population.
\end{itemize}

\subsubsection{\monthtag{December 2025}}
\begin{itemize}
    \item \textbf{Economy and Banking:} RBI 58th MPC cut repo rate to 5.25\% and retained neutral stance; SDF 5.00\%, MSF/Bank Rate 5.50\%, CRR 3.00\%, SLR 18.00\%; Governor: Sanjay Malhotra.
    \item \textbf{Economy and Banking:} RBI projected FY26 real GDP growth at 7.3\%; Q2 FY26 GDP was 8.2\%; FY26 CPI forecast cut to 2.0\%, with Q3 at 0.6\% and Q4 at 2.9\%.
    \item \textbf{Economy and Banking:} RBI announced Rs 1 trillion OMO purchases and USD 5 billion 3-year USD/INR buy-sell swap; forex reserves stood at USD 686.2 billion on 28 November, over 11 months import cover.
    \item \textbf{Economy and Banking:} UPI recorded 19 billion transactions worth Rs 24.58 lakh crore in November; IMF recognised UPI as world's largest fast payment system, with India 49\% of global real-time payments.
    \item \textbf{Economy and Banking:} PMJDY deposits reached Rs 2.75 lakh crore across 56 crore accounts; 78\% accounts rural/semi-urban and 50\% held by women.
    \item \textbf{Economy and Banking:} D-SIB list retained SBI, HDFC Bank and ICICI Bank with capital surcharges 0.80\%, 0.40\% and 0.20\%; Fino Payments Bank got in-principle nod to become SFB.
    \item \textbf{Govt Schemes:} Cabinet approved Census of India 2027 at Rs 11,718.24 crore; 16th Census and 8th after Independence; first fully digital Census with self-enumeration option and caste enumeration.
    \item \textbf{Govt Schemes:} Census 2027 phases: Houselisting and Housing Census April--September 2026; Population Enumeration February 2027; snow-bound areas in Ladakh, J\&K, HP and Uttarakhand in September 2026.
    \item \textbf{Govt Schemes:} PM SVANidhi extended to March 2030 with three-stage loans and Rs 30,000 UPI-RuPay credit card; Tex-RAMPS approved with Rs 305 crore for textile R\&D, analytics and start-ups.
    \item \textbf{Govt Schemes:} Atal Pension Yojana crossed 8.34 crore subscribers with 48\% women participation; NUDGE campaign Phase 2 by CBDT followed Phase 1 disclosure of Rs 29,208 crore foreign assets.
    \item \textbf{National:} Cabinet approved Delhi Metro Phase-V(A), 16.076 km, 13 stations, cost Rs 12,014.91 crore; corridors include R.K. Ashram Marg--Indraprastha, Aerocity--IGD T1 and Tughlakabad--Kalindi Kunj.
    \item \textbf{National:} CCEA fixed MSP for copra 2026 at Rs 12,027/quintal for milling copra and Rs 12,500/quintal for ball copra; NAFED and NCCF remain Central Nodal Agencies.
    \item \textbf{Reports/Indices:} Arton Capital Passport Index 2025: UAE rank 1 for 7th consecutive year, mobility score 179; Singapore and Spain joint 2nd, score 175; India rank 67, score 74.
    \item \textbf{Reports/Indices:} World Inequality Report 2026 by World Inequality Lab: India's top 10\% earn 58\% national income, bottom 50\% earn 15\%; top 10\% own nearly 65\% wealth and top 1\% about 40\%.
    \item \textbf{Reports/Indices:} RBI released 10th Handbook of Statistics on Indian States 2024--25; data coverage 1951 to 2024--25; 11 new state-wise tables plus new External Sector section.
    \item \textbf{International:} Russian President Vladimir Putin visited India, 4--5 December, for 23rd India-Russia Annual Summit and 25 years of Strategic Partnership; Programme 2030 set USD 100 billion bilateral trade target by 2030.
    \item \textbf{International:} India-Russia outcomes covered INSTC, Chennai--Vladivostok Maritime Route, Northern Sea Route, Kudankulam expansion, national-currency settlements, polar training and Russia joining International Big Cats Alliance.
    \item \textbf{International:} India will chair International IDEA from 3 December 2025, represented by CEC Gyanesh Kumar; Brazil transferred BRICS Presidency to India for 2026.
    \item \textbf{International:} India-Oman CEPA signed; India-Netherlands JTIC MoU signed; Japan's JERA signed first long-term LNG export deal with India's Torrent Power for 0.27 MMTPA annually for 10 years from 2027.
    \item \textbf{Environment/Health:} 20th UNESCO ICH Committee session held at Red Fort, New Delhi, 8--13 December, first time in India; Deepavali added to UNESCO Representative List; 67 new ICH elements added globally.
    \item \textbf{Environment/Health:} Chhattisgarh's Kopra Reservoir became state's first Ramsar Site on 12 December; Tamil Nadu launched Otter Conservation Programme in Cauvery basin; Gujarat declared India's newest Tiger State by NTCA.
    \item \textbf{Science-Tech:} DHRUV64, India's first indigenous 1.0 GHz 64-bit dual-core microprocessor, unveiled by C-DAC under DIR-V; Hansa-3 NG, all-composite two-seater trainer aircraft, launched.
    \item \textbf{Science-Tech:} NITI Aayog Frontier Tech Hub released Quantum Economy Roadmap targeting top-3 global quantum position by 2047; India also launched first private Navigation Excellence Centre, ACEN, in Thiruvananthapuram.
    \item \textbf{Defence:} MoD signed Rs 7,995 crore, 5-year FMS support agreement with USA for 24 MH-60R Seahawk helicopters; QUAD held first Field Training Exercise at Andersen Air Force Base, Guam, 8--12 December.
    \item \textbf{Defence:} Param Vir Dirgha Gallery inaugurated at Rashtrapati Bhavan on Vijay Diwas, 16 December; Goa Liberation Day marked 64th anniversary of 1961 Operation Vijay, India's first tri-service military operation.
    \item \textbf{Sports:} ISSF World Cup Final 2025 held at Lusail Shooting Complex, Doha, 4--9 December; China topped with 9 medals, India 2nd with 6; Simranpreet Kaur and Suruchi Singh won golds.
    \item \textbf{Sports:} World Athletics Athletes of Year 2025: Women's World Athlete Sydney McLaughlin-Levrone, Men's World Athlete Mondo Duplantis; Men's Track Emanuel Wanyonyi, Women's Field Nicola Olyslagers.
    \item \textbf{Awards/Appointments:} IFFI 2025 in Goa: Golden Peacock Best Feature Film \textit{Skin of Youth}; Satyajit Ray Lifetime Achievement Award to Rajinikanth; Country of Focus Japan, Partner Country Spain, Spotlight Country Australia.
    \item \textbf{Awards/Appointments:} National Handicrafts and Shilp Guru Awards 2023--24 presented by President Murmu at Vigyan Bhawan on 9 December; 48 awards, incl. 12 Shilp Guru and 36 National Handicrafts; 20 women awardees.
    \item \textbf{Appointments/Static GK:} Usha Janakiraman appointed RBI Executive Director; Vivek Chaturvedi appointed CBIC Chairman; Khaled El-Enany is UNESCO Director-General; UNESCO HQ Paris, established 1945.
\end{itemize}
% END SEP-DEC 2025

% BEGIN JAN-APR 2026
\subsection{January--April 2026: Monthly High-Yield Notes}

\subsubsection{\monthtag{January 2026}}
\begin{itemize}
    \item \textbf{Economy and Banking:} Government approved wage/pension revision for PSGICs, NABARD and RBI; PSGIC wage revision effective 1 August 2022 with 12.41\% wage-bill hike and 14\% increase on Basic Pay + DA; 43,247 employees covered.
    \item \textbf{Economy and Banking:} PSGIC NPS contribution raised from 10\% to 14\% for employees joining after 1 April 2010; uniform family pension rate fixed at 30\%; total PSGIC financial outgo Rs 8,170.30 crore.
    \item \textbf{Economy and Banking:} NABARD pay revision effective 1 November 2022 gives about 20\% hike to Group A, B and C employees; RBI pension/family pension revision effective 1 November 2022 gives 10\% increase, benefiting 30,769 retirees/family pensioners.
    \item \textbf{Economy and Banking:} Commerce Ministry launched MSME export interventions worth Rs 7,295 crore: Rs 5,181 crore interest subvention and Rs 2,114 crore collateral guarantee; guarantee cover 85\% for micro/small exporters and 65\% for medium exporters.
    \item \textbf{Economy and Banking:} World Bank upgraded India's FY26 GDP growth forecast to 7.2\% from 6.3\%; India-Germany bilateral trade crossed USD 50 billion with 19 strategic agreements.
    \item \textbf{Govt Schemes:} Startup India completed 10 years; PM launched 5th National Startup Awards across 20 categories focused on deep-tech innovation and Tier-I/Tier-III city startups.
    \item \textbf{Govt Schemes:} Cabinet approved Rs 5,000 crore equity support to SIDBI in three tranches during FY26--FY28 to expand MSME credit and employment generation.
    \item \textbf{Govt Schemes:} Market Access Support Scheme announced with Rs 4,531 crore outlay for FY26--FY31 to support MSMEs and first-time exporters through fairs, exhibitions, buyer-seller meets and trade delegations.
    \item \textbf{Govt Schemes:} Punjab introduced Mukhya Mantri Sehat Yojana: free cashless healthcare up to Rs 10 lakh per family annually for 65 lakh families.
    \item \textbf{Govt Schemes:} Punjab launched Mukh Mantri Mawan Dhiyan Satkar Yojana in budget context: Rs 1,000/month to women and Rs 1,500/month to Dalit women, with Rs 9,300 crore allocation.
    \item \textbf{National/Infrastructure:} Andhra Pradesh Cabinet approved Dagadarthi Greenfield Airport in Nellore as state's 8th airport with 1.4 million passenger capacity in Phase-I.
    \item \textbf{National/Infrastructure:} AP CM laid foundation for world's largest green hydrogen and ammonia project at Kakinada: USD 10 billion investment, 1.5 MTPA capacity across 3 phases.
    \item \textbf{National/Infrastructure:} Assam received India's first Sleeper Vande Bharat Express on Howrah--Guwahati route; PM also laid foundation for Rs 6,957 crore Kaziranga elevated corridor.
    \item \textbf{National/Infrastructure:} GAIL completed 694-km Mumbai--Nagpur Natural Gas Pipeline, 96\% along Samruddhi Mahamarg corridor under PM GatiShakti model.
    \item \textbf{Reports/Indices:} Henley Passport Index 2026: Singapore rank 1 with 192 visa-free destinations; Japan and South Korea joint rank 2 with 188; India rank 80 with access to 55 destinations.
    \item \textbf{Reports/Indices:} Bottom passports in Henley 2026: Afghanistan rank 101 with 24 destinations, Syria rank 100 with 26, Iraq rank 99 with 29, Pakistan/Yemen rank 98 with 31.
    \item \textbf{Reports/Indices:} Export Preparedness Index 2024 by NITI Aayog: Maharashtra topped large states with score 68.01, followed by Tamil Nadu 64.41 and Gujarat 64.02.
    \item \textbf{International:} UAE President Sheikh Mohamed bin Zayed Al Nahyan visited India; outcomes included LoI on Dholera SIR, IN-SPACe--UAE Space Agency LoI, strategic defence partnership LoI and HPCL--ADNOC Gas SPA for 0.5 MMTPA LNG for 10 years from 2028.
    \item \textbf{International:} India-UAE set bilateral trade doubling target of USD 200 billion by 2032; cooperation areas included AI supercomputing cluster, civil nuclear cooperation, GIFT City, Digital/Data Embassies and House of India in Abu Dhabi.
    \item \textbf{International/Summits:} WEF Annual Meeting 2026 held at Davos, Switzerland; theme ``A Spirit of Dialogue''; nearly 3,000 global leaders participated, with focus on geopolitics, AI, trade, reskilling and resilient supply chains.
    \item \textbf{Environment/Health:} Project Dolphin second nationwide survey launched, covering major river systems including Ganga Basin to estimate dolphin populations and monitor freshwater ecosystems.
    \item \textbf{Environment/Health:} India non-fossil fuel capacity reached 266.78 GW in 2025, up 22.6\%; solar capacity 131.81 GW; India added 48 GW renewable energy in 2025.
    \item \textbf{Environment/Health:} GoI approved Nutrient-Based Subsidy rates for Rabi 2025--26 with budget requirement Rs 37,952.29 crore; DAP subsidy raised to Rs 29,805/MT.
    \item \textbf{Science-Tech:} PSLV-C62/EOS-N1 mission launched on 12 January 2026 from Sriharikota; anomaly occurred in PS3 stage; mission was PSLV's 64th flight and 105th launch from the spaceport.
    \item \textbf{Science-Tech:} PSLV-C62 carried EOS-N1 and 15 co-passenger satellites; DRDO's Anvesha surveillance satellite was onboard; Dhruva Space contributed 7 satellites, a major private-sector milestone.
    \item \textbf{Science-Tech:} MeitY approved 22 proposals worth Rs 41,863 crore under Electronics Components Manufacturing Scheme; cumulative approvals reached 46 projects with Rs 56,567 crore investment.
    \item \textbf{Science-Tech:} Tamil Nadu launched India's first state-backed Quantum Computing Programme with German firm XeedQ GmbH; Azista Space laid foundation for Palmnaro private satellite manufacturing facility at Sanand, Gujarat.
    \item \textbf{Defence/Security:} Indian Army declared 2026 as ``Year of Networking \& Data Centricity'' to integrate soldiers, command centres, sensors and weapon platforms through secure digital connectivity.
    \item \textbf{Defence/Security:} Army launched ``Radio Sangam'', first community radio station along the LoC, at Keri village in Rajouri, J\&K, broadcasting on 88.8 FM.
    \item \textbf{Defence/Security:} India and Pakistan exchanged nuclear-installation lists on 1 January for the 35th consecutive time under the 1988 agreement; prisoner lists were also exchanged.
    \item \textbf{Awards/Appointments:} Rakesh Aggarwal, IPS 1994 batch, appointed Director General of NIA with tenure till August 2028.
    \item \textbf{Awards:} Golden Globes 2026: Best Film Drama -- \textit{Hamnet}; Best Film Musical/Comedy -- \textit{One Battle After Another}; Best Director and Screenplay -- Paul Thomas Anderson.
    \item \textbf{Awards:} Golden Globes acting winners included Wagner Moura for \textit{The Secret Agent}, Jessie Buckley for \textit{Hamnet}, Timothee Chalamet for \textit{Marty Supreme} and Rose Byrne for \textit{If I Had Legs I'd Kick You}.
    \item \textbf{Sports/Static GK:} International Kite Festival 2026 inaugurated at Sabarmati Riverfront with German Chancellor Friedrich Merz as Chief Guest; linked to Uttarayan celebrations in Gujarat.
    \item \textbf{Static GK:} BRICS Chairship 2026 logo and theme unveiled: ``Building for Resilience, Innovation, Cooperation and Sustainability''; India's chairship is its 4th and coincides with BRICS' 20th anniversary.
\end{itemize}

\subsubsection{\monthtag{February 2026}}
\begin{itemize}
    \item \textbf{Economy and Banking:} Union Budget 2026--27 presented by Finance Minister Nirmala Sitharaman, first Budget prepared in Kartavya Bhawan; fiscal deficit 4.3\% of GDP, debt-to-GDP 55.6\%, total expenditure Rs 53.5 lakh crore.
    \item \textbf{Economy and Banking:} Budget estimates: non-debt receipts Rs 36.5 lakh crore, Centre's net tax receipts Rs 28.7 lakh crore, gross market borrowings Rs 17.2 lakh crore, public capital expenditure Rs 12.2 lakh crore.
    \item \textbf{Economy and Banking:} RBI 59th MPC, 4--6 February 2026, kept repo rate unchanged at 5.25\%; SDF 5.00\%, MSF/Bank Rate 5.50\%; stance neutral; Governor Sanjay Malhotra.
    \item \textbf{Economy and Banking:} RBI February projections: 2025--26 GDP growth 7.4\%, Q1 FY27 6.9\%, Q2 FY27 7.0\%; CPI projection 2025--26 at 2.1\%, Q4 FY26 3.2\%, Q1 FY27 4.0\%, Q2 FY27 4.2\%.
    \item \textbf{Economy and Banking:} 16th Finance Commission retained states' tax devolution at 41\% for 2026--31; introduced 10\% weight for State GDP contribution and recommended zero Revenue Deficit Grants.
    \item \textbf{Economy and Banking:} 16th FC earmarked Rs 7.91 trillion for rural/urban local bodies with 60:40 rural-urban split; Rs 2.04 trillion for State Disaster Response and Mitigation Funds and Rs 79,000 crore for national disaster funds.
    \item \textbf{Govt Schemes:} Budget announced Biopharma SHAKTI Initiative with Rs 10,000 crore outlay over 5 years; 3 new NIPERs, upgrade of 7 existing NIPERs and 1,000+ clinical trial sites.
    \item \textbf{Govt Schemes:} India Semiconductor Mission 2.0 to focus on equipment, materials and full-stack Indian IP; Electronics Components Manufacturing Scheme outlay increased to Rs 40,000 crore.
    \item \textbf{Govt Schemes:} Startup India Fund of Funds 2.0 approved with Rs 10,000 crore corpus to support deep-tech, AI, robotics, semiconductor and clean-energy startups via AIFs.
    \item \textbf{Govt Schemes:} Urban Challenge Fund approved with Rs 1 lakh crore Central Assistance for FY26--31; expected to catalyse Rs 4 lakh crore investment; at least 50\% market-based financing mandated.
    \item \textbf{Govt Schemes:} PM-SETU Programme: Rs 60,000 crore allocation to upgrade 1,000 ITIs with modern equipment, global-standard curricula and industry linkages.
    \item \textbf{Govt Schemes:} White Revolution 2.0 launched by Ministry of Cooperation to increase milk procurement by 50\% over 5 years; E20 petrol mandate starts nationwide from 1 April 2026.
    \item \textbf{National/Infrastructure:} New PMO named ``Seva Teerth'' dedicated on 13 February; North and South Blocks to become ``Yuge Yugeen Bharat National Museum''.
    \item \textbf{National/Infrastructure:} CCEA approved India's first underwater road-cum-rail tunnel under Brahmaputra: 33.7 km Gohpur--Numaligarh corridor, 15.79 km twin-tube tunnel, cost Rs 18,662 crore.
    \item \textbf{National/Infrastructure:} CCEA approved three railway projects worth Rs 9,072 crore: Gondia--Jabalpur doubling, Punarakh--Kiul 3rd/4th line and Gamharia--Chandil 3rd/4th line; 307 km addition, 52 MTPA freight capacity.
    \item \textbf{National/Infrastructure:} New Civil Enclave at Srinagar International Airport approved at Rs 1,677 crore; 73.18 acres, 71,500 sq m terminal, 10 MPPA capacity.
    \item \textbf{Reports/Indices:} India ranked 45th in Network Readiness Index 2025; USA topped. India ranked 91st in Corruption Perceptions Index 2025 with score 39; Denmark retained rank 1.
    \item \textbf{Reports/Indices:} India ranked 131st in Women, Peace and Security Index; Denmark led. India ranked 3rd globally in billionaires with 188, after USA-led list.
    \item \textbf{Reports/Indices:} India ranked 2nd globally in LEED-certified green building space in USGBC LEED List 2025; China topped.
    \item \textbf{International:} India hosted 2nd India-Arab Foreign Ministers Meeting in New Delhi with 22 Arab League countries; New Delhi Declaration adopted.
    \item \textbf{International:} India-USA trade deal reduced tariffs from 25\% to 18\%, projected to narrow India's CAD by 0.25\% of GDP; India and China trade reached USD 155.6 billion in 2025, Indian exports up 9.7\%.
    \item \textbf{International:} India chaired first BRICS Sherpas' Meeting 2026 in New Delhi under theme ``Building for Resilience, Innovation, Cooperation and Sustainability''.
    \item \textbf{International/Sports:} Milano Cortina Winter Olympics 2026 held in Italy from 6--22 February with 16 winter sports; India participated with two athletes.
    \item \textbf{Environment/Health:} WHO prequalified novel Oral Polio Vaccine nOPV2 developed by Biological E Limited, Hyderabad.
    \item \textbf{Environment/Health:} Supreme Court declared Menstrual Hygiene Management a Fundamental Right under Article 21; Alwar, Rajasthan became India's first fully insured district under ``Insurance for All by 2047''.
    \item \textbf{Science-Tech:} India AI Impact Summit 2026 held at Bharat Mandapam, New Delhi; theme ``Sarvajan Hitaya, Sarvajan Sukhaya''; participants included 100+ countries, 20+ heads of state/government, 60 ministers and 500+ AI leaders/CEOs.
    \item \textbf{Science-Tech:} PM unveiled M.A.N.A.V. AI framework: Moral and Ethical Systems, Accountable Governance, National Sovereignty, Accessible and Inclusive AI, Valid and Legitimate AI systems.
    \item \textbf{Science-Tech:} India AI Impact Summit pillars: People, Planet and Progress; working groups covered economic growth, democratizing AI resources, social inclusion, safe/trusted AI, human capital, science and resilience.
    \item \textbf{Science-Tech:} NSO launched Model Context Protocol server on eSankhyiki Portal for AI-enabled access to CPI, IIP and official datasets; Qualcomm's 2nm semiconductor chip design launched in Bengaluru.
    \item \textbf{Defence/Security:} Ministry of Home Affairs launched PRAHAAR, India's National Counter-Terrorism Policy, based on seven pillars; Delhi Police launched Operation Shastra to monitor 6,000+ social-media profiles.
    \item \textbf{Defence/Security:} Yantra India Limited granted Miniratna Category-I status, allowing independent capital expenditure approvals up to Rs 500 crore.
    \item \textbf{Awards:} Grammy Awards 2026: Song of the Year -- Billie Eilish, ``Wildflower''; Best New Artist -- Olivia Dean; Producer of the Year -- Cirkut; Songwriter of the Year -- Amy Allen.
    \item \textbf{Awards:} BAFTA 2026: Best Film -- \textit{One Battle After Another}; Outstanding British Film -- \textit{Hamnet}; Director -- Paul Thomas Anderson; Leading Actress -- Jessie Buckley; Leading Actor -- Robert Aramayo.
    \item \textbf{Sports:} WPL 2026: Royal Challengers Bengaluru beat Delhi Capitals by 6 wickets at Vadodara; chased 204, highest successful WPL/IPL final chase; Smriti Mandhana Player of Match with 87 off 41 balls.
    \item \textbf{Sports:} WPL 2026 awards: Winner RCB Rs 6 crore, Runner-up DC Rs 3 crore; Orange Cap Smriti Mandhana, Purple Cap and MVP Sophie Devine, Emerging Player Nandini Sharma.
    \item \textbf{Sports:} Australian Open 2026: Men's Singles -- Carlos Alcaraz; Women's Singles -- Elena Rybakina; Men's Doubles -- Neal Skupski/Christian Harrison; Women's Doubles -- Elise Mertens/Zhang Shuai.
    \item \textbf{Static GK:} Arunachal Pradesh and Mizoram celebrate Statehood Day on 20 February; both attained statehood in 1987.
\end{itemize}

\subsubsection{\monthtag{March 2026}}
\begin{itemize}
    \item \textbf{Economy and Banking:} State budgets: Haryana Rs 2.23 lakh crore, Chhattisgarh Rs 1.72 lakh crore, Goa Rs 30,195 crore, Punjab Rs 2,60,437 crore, Uttarakhand Rs 1,11,703 crore.
    \item \textbf{Economy and Banking:} Karnataka Budget 2026--27 outlay Rs 4.48 lakh crore; Maharashtra Budget Rs 7,69,467 crore with fiscal deficit 2.7\% GSDP; Telangana Budget Rs 3,24,234 crore with GSDP growth estimate 10.7\%.
    \item \textbf{Economy and Banking:} Delhi Budget 2026--27 outlay Rs 1,03,700 crore; called a Green Budget with 21\% allocation for green initiatives; projected revenue surplus Rs 9,092 crore.
    \item \textbf{Economy and Banking:} Bharat Electricity Summit 2026 held at Yashobhoomi, Dwarka, New Delhi; theme ``Electrifying Growth. Empowering Sustainability. Connecting Globally''; Indian Carbon Market Portal launched.
    \item \textbf{Govt Schemes:} PM released 22nd instalment of PM-KISAN during Assam visit, transferring Rs 2,000 each to 9.3 crore farmers.
    \item \textbf{Govt Schemes:} Delhi launched four women-centric schemes: Lakhpati Bitiya, Saheli Pink Smart Card, Free LPG Cylinder Scheme and Meri Poonji Mera Adhikaar.
    \item \textbf{Govt Schemes:} Chhattisgarh launched Gaudham Yojana to establish 1,460 cattle shelters; 10 per Development Block; daily feed assistance Rs 10--35 per cattle.
    \item \textbf{Govt Schemes:} Punjab Sikhya Kranti 2.0 launched with Rs 3,500 crore investment over 6 years, including Rs 2,500 crore World Bank loan and Rs 1,000 crore state share.
    \item \textbf{National/Infrastructure:} PM inaugurated India's first fully operational Ring Metro corridor in Delhi by completing 12.3 km Majlis Park--Maujpur-Babarpur section of Pink Line; Pink Line length 71.56 km.
    \item \textbf{National/Infrastructure:} PM inaugurated 9.9 km Magenta Line extension from Deepali Chowk to Majlis Park with 7 stations; Delhi Metro Phase-V(A) corridors also launched.
    \item \textbf{National/Infrastructure:} PM visited Assam and West Bengal on 13--14 March; projects worth over Rs 66,280 crore announced/inaugurated: Rs 47,600 crore in Assam and Rs 18,680 crore in West Bengal.
    \item \textbf{National/Infrastructure:} Noida International Airport Phase-I at Jewar planned for inauguration on 28 March; cost Rs 11,200 crore, 12 million annual passenger capacity, 3,900 m runway, net-zero design.
    \item \textbf{National/Legal:} Supreme Court ruled SC status is limited to Hindus, Sikhs and Buddhists under Constitution (Scheduled Castes) Order, 1950; conversion to Christianity removes SC status and related SC/ST Act protection.
    \item \textbf{National/Legal:} Supreme Court-appointed panel chaired by Justice Asha Menon urged withdrawal of Transgender Persons (Protection of Rights) Amendment Bill 2026, citing risks to self-identification rights under NALSA judgment.
    \item \textbf{Reports/Indices:} World Happiness Report 2026: Finland rank 1 for 9th consecutive year; Denmark 2, Iceland 3, Costa Rica 4; India rank 116 of 147, behind China rank 60 and Pakistan rank 104.
    \item \textbf{Reports/Indices:} Top 10 richest people, March 2026: Elon Musk rank 1 with USD 841.4 billion; Larry Page rank 2 USD 250.5 billion; Sergey Brin rank 3 USD 231.2 billion.
    \item \textbf{Reports/Indices:} Odisha retained rank 1 in Fiscal Health Index 2026 with score 73.1.
    \item \textbf{International:} India-Germany launched EUR 20 million climate resilience programme under International Climate Initiative for Himalayas, Western Ghats and North-East India.
    \item \textbf{International/Science-Tech:} India AI Impact Summit 2026 outcome: declaration endorsed by 92 countries/international organisations; over 2.5 lakh responsible AI pledges in 24 hours created a Guinness World Record.
    \item \textbf{International/Science-Tech:} IndiaAI Mission compute capacity: 38,000+ GPUs provisioned and 20,000 more to be added; summit generated over USD 200 billion AI investment commitments.
    \item \textbf{Science-Tech:} Reliance announced USD 110 billion AI infrastructure plan over seven years; Adani Enterprises announced USD 100 billion by 2035; Google announced USD 15 billion AI hub in Visakhapatnam.
    \item \textbf{Science-Tech:} Census 2027 digital tools and mascots soft-launched: Houselisting Block Creator Web App, House Listing Officer Mobile App, Self-Enumeration Portal, CMMS Portal; mascots Pragati and Vikas.
    \item \textbf{Science-Tech:} New Delhi deployed India's first microalgae-based PureAir Tower along Aerocity highway; one tower's air-purification effect compared to 15 mature trees.
    \item \textbf{Science-Tech:} NDSA launched NETRA AI-enabled dam-safety platform with Indian Navy and Rashtriya Bandh Suraksha Darpan by C-DAC for dam-break analysis.
    \item \textbf{Environment/Health:} Tamil Nadu launched first Dark Sky Park at Kolli Hills, Namakkal district, with Rs 1 crore investment, 3 telescopes and annual operation from 15 January to 15 June.
    \item \textbf{Environment/Health:} Varanasi set Guinness World Record by planting 2,51,446 saplings in one hour at Sujabad Domari using Miyawaki technique, beating China's 2018 record of 1,53,981.
    \item \textbf{Environment/Health:} Project HANUMAN launched in Andhra Pradesh for human-wildlife conflict: 100 GPS-enabled vehicles, 4 rescue/treatment centres and death compensation raised from Rs 5 lakh to Rs 10 lakh.
    \item \textbf{Defence/Security:} Uttarakhand ranked 1st nationally in implementing new criminal laws through ICJS 2.0, scoring 93.46 on CCTNS Progress Dashboard.
    \item \textbf{Defence/Security:} Maharashtra passed Freedom of Religion Bill 2026: up to 7 years imprisonment for forced conversions through marriage/fraud/coercion; up to 10 years for repeat offenders.
    \item \textbf{Awards:} Oscars 2026 at Dolby Theatre, Los Angeles, hosted by Conan O'Brien; Best Picture -- \textit{One Battle After Another}; most wins 6; most nominations -- \textit{Sinners} with 16.
    \item \textbf{Awards:} Oscars major winners: Best Actor Michael B. Jordan for \textit{Sinners}; Best Actress Jessie Buckley for \textit{Hamnet}; Best Director Paul Thomas Anderson for \textit{One Battle After Another}; Best International Feature \textit{Sentimental Value}.
    \item \textbf{Awards:} Sahitya Akademi Award 2025 ceremony on 31 March in New Delhi; prize Rs 1,00,000, shawl and engraved copper plaque; English winner Navtej Sarna for novel \textit{Crimson Spring}.
    \item \textbf{Awards:} Ramnath Goenka Awards 2026 notable winners: Avadhesh Akodia (Print Hindi), Sarvapriya Sangwan (Broadcast/Digital Hindi), Jayashree Nandi and Tannu Jain (Environment, Science \& Technology Reporting).
    \item \textbf{Appointments:} Sadhvi Niranjan Jyoti assumed charge as Chairperson of National Commission for Backward Classes; Kiran Umesh Mahalle assumed charge as NCBC Member.
    \item \textbf{Sports:} ICC Men's T20 World Cup 2026 final at Narendra Modi Stadium, Ahmedabad: India beat New Zealand by 96 runs; India 255/5, New Zealand 159 all out; Sanju Samson 89, Jasprit Bumrah 4/15.
    \item \textbf{Sports:} India secured back-to-back T20 World Cup titles; other final performers: Abhishek Sharma 52 off 21, Ishan Kishan 54 off 25, Axar Patel 3/27.
    \item \textbf{Important Days:} Janaushadhi Saptah 2026 observed 1--7 March; theme ``Jan Aushadhi -- affordable and trustworthy, good health with savings''.
    \item \textbf{Important Days:} National Science Day 2026 observed on 28 February at Vigyan Bhawan; theme ``Women in Science: Catalysing Viksit Bharat''; marks Raman Effect discovery by C.V. Raman in 1928.
    \item \textbf{Static GK:} Kalinjar Fort Hill, Banda district, UP, declared National Geo-Heritage Site by GSI; features Eparchaean Unconformity with Bundelkhand granite and Kaimur sandstone.
\end{itemize}

\subsubsection{\monthtag{April 2026}}
\begin{itemize}
    \item \textbf{Economy and Banking:} RBI 60th MPC, 6--8 April 2026, kept repo rate unchanged at 5.25\%; SDF 5.00\%, MSF/Bank Rate 5.50\%; stance neutral; next MPC scheduled 3--5 June 2026.
    \item \textbf{Economy and Banking:} RBI April projections: GDP growth 2025--26 estimated 7.6\%, 2026--27 projected 6.9\%; CPI inflation February 2026 at 3.2\%, 2026--27 projected 4.6\%.
    \item \textbf{Economy and Banking:} Global exports of goods reached USD 26.3 trillion in 2025, up 7\%, as per WTO; top 10 exporters accounted for 49.6\% of global exports.
    \item \textbf{Economy and Banking:} Global export leaders 2025: China rank 1 USD 3,771,842 million; USA rank 2 USD 2,185,220 million; Germany rank 3 USD 1,764,188 million; India rank 19 USD 445,278 million, 1.7\% share.
    \item \textbf{Economy and Banking:} PM Mudra Yojana completed 11 years since launch on 8 April 2015; 57.79 crore loans sanctioned worth Rs 40.07 lakh crore; about 67\% beneficiaries women and 51\% SC/ST/OBC.
    \item \textbf{Economy and Banking:} PMMY loan categories: Shishu up to Rs 50,000; Kishor Rs 50,000--Rs 5 lakh; Tarun Rs 5 lakh--Rs 10 lakh; Tarun Plus Rs 10 lakh--Rs 20 lakh.
    \item \textbf{Govt Schemes:} Cabinet approved NBS rates for Kharif 2026 on P\&K fertilizers with estimated budget Rs 41,533.81 crore.
    \item \textbf{Govt Schemes:} Cabinet approved extension of PMGSY-III till March 2028 with outlay Rs 83,977 crore for rural connectivity to markets, schools and hospitals.
    \item \textbf{Govt Schemes:} DA/DR raised by 2 percentage points to 60\% of basic pay/pension from January 2026; benefits 50.46 lakh employees and 68.27 lakh pensioners; annual impact Rs 6,791.24 crore.
    \item \textbf{Govt Schemes:} Bharat Maritime Insurance Pool approved with sovereign guarantee Rs 12,980 crore to insure Indian vessels against cargo, war and liability risks.
    \item \textbf{National/Infrastructure:} Census 2027 Phase 1, Houselisting and Housing Census, began April 2026; first fully digital Census with mobile apps, self-enumeration in 16 languages and caste enumeration in Population Enumeration phase.
    \item \textbf{National/Infrastructure:} Census 2027 is India's 16th Census and 8th after Independence; Phase 1 runs April--September 2026; Population Enumeration in February 2027.
    \item \textbf{National/Infrastructure:} PM inaugurated Kaynes Semicon plant at Sanand, Gujarat, on 31 March 2026; supports Intelligent Power Modules for EVs, industrial automation and electronics.
    \item \textbf{National/Infrastructure:} India's semiconductor projects worth over Rs 1.6 lakh crore were underway; Semiconductor Mission 2.0 targets 85,000 semiconductor professionals and supports Chips to Startup programme with 400+ universities/innovators.
    \item \textbf{National/Infrastructure:} Cabinet approved Kamala Hydro Electric Project, Arunachal Pradesh: 1,720 MW, Rs 26,069.50 crore, expected 6,870 MU generation.
    \item \textbf{National/Infrastructure:} Cabinet approved Kalai-II Hydro Electric Project on Lohit River, Arunachal Pradesh: 1,200 MW, Rs 14,105.83 crore, expected 4,852.95 MU annually.
    \item \textbf{National/Infrastructure:} Cabinet approved railway multitracking projects worth Rs 24,815 crore in UP and Andhra Pradesh, adding 601 km on Ghaziabad--Sitapur and Rajahmundry--Visakhapatnam routes.
    \item \textbf{National/Legal:} Supreme Court ruled right to vote and right to contest elections are statutory rights, not fundamental rights; case related to Rajasthan cooperative-society elections.
    \item \textbf{National/Legal:} Supreme Court cited Representation of the People Acts, 1950 and 1951, as governing electoral rights; cooperative societies generally not ``State'' under Article 12.
    \item \textbf{National/Political:} Women's Reservation Bill 2026, Constitutional 131st Amendment Bill, failed in Lok Sabha despite 298 votes in favour and 230 against because required 2/3 majority threshold was 360.
    \item \textbf{National/Political:} Amaravati declared sole capital of Andhra Pradesh by Gazette notification dated 6 April 2026, ending three-capital model.
    \item \textbf{National/Appointments:} Samrat Choudhary sworn in as Bihar Chief Minister after Nitish Kumar resigned and moved to Rajya Sabha; Ashwini Bhide appointed first woman Municipal Commissioner of BMC.
    \item \textbf{Reports/Indices:} TIME100 2026, 23rd annual list, featured Indian names Sundar Pichai, Vikas Khanna and Ranbir Kapoor; youngest Alysa Liu, age 20, oldest Dolores Huerta, age 96.
    \item \textbf{Reports/Indices:} TIME100 Summit held 22 April in New York City; Gala on 23 April hosted by Nikki Glaser with performances by Luke Combs and Coco Jones.
    \item \textbf{International/Space:} NASA launched Artemis II on 1 April 2026 from Kennedy Space Center at 6:35 p.m. ET; first crewed Moon journey since 1972, 10-day Orion mission around the Moon.
    \item \textbf{International/Space:} Artemis II crew: Reid Wiseman, Victor Glover, Christina Koch and Jeremy Hansen; Glover is first Black astronaut to deep space, Koch first woman to deep space, Hansen first non-American on a Moon mission.
    \item \textbf{Environment/Health:} Goldman Environmental Prize 2026, the ``Green Nobel'', awarded first all-women cohort; each winner receives USD 200,000.
    \item \textbf{Environment/Health:} Goldman 2026 winners: Iroro Tanshi (Nigeria), Borim Kim (South Korea), Sarah Finch (UK), Theonila Roka Matbob (Papua New Guinea), Alannah Acaq Hurley (USA), Yuvelis Morales Blanco (Colombia).
    \item \textbf{Environment/Health:} Great Indian Bustard chick born in Gujarat after nearly 10 years using ``Jumpstart Approach''; Solid Waste Management Rules 2026 mandated four-way waste segregation.
    \item \textbf{Science-Tech:} Gujarat Police launched NARIT AI, India's first AI-based narcotics investigation tool under NDPS Act, 1985; uses Retrieval-Augmented Generation to analyse FIRs and legal gaps.
    \item \textbf{Science-Tech:} Gujarat High Court banned AI use in judicial decision-making and judgment drafting, while permitting AI for legal research and administrative work.
    \item \textbf{Science-Tech:} Ministry of Education introduced AI \& Computational Thinking curriculum for Classes 3--8; Election Commission launched KYC feature on ECINet for candidate criminal records, assets and education details.
    \item \textbf{Science-Tech:} India laid foundation of first 3D semiconductor packaging unit in Odisha; CBI launched AI chatbot ``Abhay'' to help citizens verify official notices and prevent cyber fraud.
    \item \textbf{Defence:} Indian Navy inducted INS Aridaman, third nuclear-powered ballistic missile submarine of Arihant class; India has three operational SSBNs at sea for first time.
    \item \textbf{Defence:} INS Aridaman features about 7,000-tonne displacement, more VLS tubes, K-15 and K-4 nuclear-capable missiles; K-4 range up to 3,500 km.
    \item \textbf{Defence/Static GK:} Nuclear triad means ability to launch nuclear weapons from land-based missiles, aircraft and sea-based submarines; strengthens second-strike capability.
    \item \textbf{Awards:} Chetak Screen Awards 2026: Best Film -- \textit{Homebound}; Best Director -- Aditya Dhar for \textit{Dhurandhar}; Best Actor -- Ranveer Singh; Best Actress -- Yami Gautam.
    \item \textbf{Sports:} Laureus World Sports Awards 2026 held in Madrid; Sportsman -- Carlos Alcaraz; Sportswoman -- Aryna Sabalenka; Team -- Paris Saint-Germain; Breakthrough -- Lando Norris.
    \item \textbf{Sports:} Laureus 2026: Comeback -- Rory McIlroy; Sportsperson with Disability -- Gabriel Araujo; Action Sportsperson -- Chloe Kim; Young Sportsperson -- Lamine Yamal; Lifetime Achievement -- Nadia Comaneci.
    \item \textbf{Summits/Important Days:} Samrat Samprati Museum inaugurated by PM at Shri Mahavir Jain Aradhana Kendra, Koba village, Gandhinagar, on Mahavir Jayanti; museum has nearly 2,000 Jain artefacts in 7 galleries.
    \item \textbf{Static GK:} Samrat Samprati ruled c. 224--215 BCE and is remembered as a major patron of Jainism and ahimsa; often compared with Ashoka for religious patronage.
\end{itemize}
% END JAN-APR 2026

\section{Topic-Wise Master Revision}

% BEGIN TOPIC MASTER
\subsection{Economy/Banking}
\begin{itemize}
    \fact{May 2025}{IMF WEO placed India as 4th-largest economy in 2025: India USD 4.187 trillion vs Japan USD 4.186 trillion; growth 6.2\% in 2025 and 6.3\% in 2026.}
    \fact{Jun 2025}{RBI 55th MPC cut repo by 50 bps to 5.50\%; SDF 5.25\%, MSF/Bank Rate 5.75\%; stance neutral; FY26 GDP 6.5\%, CPI 3.7\%.}
    \fact{Aug--Oct 2025}{Repo held at 5.50\%; August CPI forecast 3.1\%, October FY26 GDP forecast raised to 6.8\% and CPI cut to 2.6\%; CRR 3\% in October.}
    \fact{Dec 2025--Apr 2026}{RBI 58th MPC cut repo to 5.25\%; April 2026 kept repo 5.25\%, SDF 5.00\%, MSF/Bank Rate 5.50\%; FY27 GDP projected 6.9\%, CPI 4.6\%.}
    \fact{Feb 2026}{Union Budget 2026--27: fiscal deficit 4.3\% of GDP, debt-to-GDP 55.6\%, total expenditure Rs 53.5 lakh crore, public capex Rs 12.2 lakh crore.}
    \fact{Feb 2026}{16th Finance Commission retained states' tax devolution at 41\% for 2026--31; Rs 7.91 trillion for local bodies; zero Revenue Deficit Grants.}
    \fact{Sep--Dec 2025}{GST slabs simplified mainly to 5\% and 18\%; GST collections Rs 1.86 trillion in August and Rs 1.89 lakh crore in September; December forex reserves USD 686.2 billion.}
    \fact{Nov--Dec 2025}{India became world's 3rd-largest equity market above USD 5 trillion; UPI touched Rs 27.28 lakh crore/20.7 billion transactions in October and Rs 24.58 lakh crore/19 billion in November.}
    \fact{Dec 2025}{PMJDY deposits Rs 2.75 lakh crore across 56 crore accounts; D-SIBs: SBI 0.80\%, HDFC Bank 0.40\%, ICICI Bank 0.20\% capital surcharge.}
    \fact{Apr 2026}{Global goods exports 2025 USD 26.3 trillion; China rank 1, USA 2, Germany 3, India rank 19 with USD 445,278 million and 1.7\% share.}
    \fact{Apr 2026}{PM Mudra Yojana completed 11 years: 57.79 crore loans worth Rs 40.07 lakh crore; Shishu up to Rs 50,000, Kishor Rs 50,000--5 lakh, Tarun Rs 5--10 lakh, Tarun Plus Rs 10--20 lakh.}
\end{itemize}

\subsection{Schemes}
\begin{itemize}
    \fact{May/Oct 2025}{PM SETU/ITI upgradation: Rs 60,000 crore, 1,000 ITIs, 5 NSTIs, 5 Centres of Excellence, 20 lakh youth over 5 years.}
    \fact{May 2025}{Modified Interest Subvention Scheme FY26: KCC loans up to Rs 3 lakh at 7\%; 1.5\% bank subvention, 3\% prompt repayment incentive, effective 4\%; 7.75 crore farmers.}
    \fact{Jul 2025}{Employment Linked Incentive Scheme: Rs 99,446 crore, 3.5 crore jobs in 2 years; first-time employees get one month's wage up to Rs 15,000.}
    \fact{Jul 2025}{RDI Scheme: Rs 1 lakh crore corpus via ANRF; PM Dhan-Dhaanya Krishi Yojana for 100 districts, 36 schemes and 11 departments.}
    \fact{Aug/Dec 2025}{PM SVANidhi extended to March 2030: Rs 7,332 crore outlay, 1.15 crore vendors, loan slabs Rs 15,000/Rs 25,000/Rs 50,000, UPI-RuPay credit card.}
    \fact{Oct--Nov 2025}{Mission for Aatmanirbharta in Pulses Rs 11,440 crore; Export Promotion Mission Rs 25,060 crore; REPM scheme Rs 7,280 crore for 6,000 MTPA ecosystem.}
    \fact{Dec 2025--Apr 2026}{Census 2027 approved at Rs 11,718.24 crore; 16th Census, 8th after Independence, first fully digital with self-enumeration and caste enumeration.}
    \fact{Feb 2026}{Biopharma SHAKTI Rs 10,000 crore; Startup India Fund of Funds 2.0 Rs 10,000 crore; Urban Challenge Fund Rs 1 lakh crore to catalyse Rs 4 lakh crore.}
    \fact{Feb 2026}{White Revolution 2.0 targets 50\% rise in milk procurement in 5 years; E20 petrol mandate starts nationwide from 1 April 2026.}
    \fact{Apr 2026}{PMGSY-III extended till March 2028 with Rs 83,977 crore; Bharat Maritime Insurance Pool sovereign guarantee Rs 12,980 crore.}
\end{itemize}

\subsection{Reports/Indices}
\begin{itemize}
    \fact{Jun 2025}{QS World University Rankings 2026: MIT rank 1; IIT Delhi top Indian at 123, IIT Bombay 129, IIT Madras 180.}
    \fact{Jun 2025}{Global Peace Index 2025: Iceland rank 1, Ireland 2, New Zealand 3; India rank 115, Russia least peaceful at 163.}
    \fact{Jul 2025}{UN SDG Index 2025: Finland rank 1, Sweden 2, Denmark 3; India rank 99, first time in top 100; South Sudan rank 167.}
    \fact{Sep 2025}{Global Innovation Index 2025: India rank 38, Switzerland rank 1; NIRF 2025 overall top 3 IIT Madras, IISc Bengaluru, IIT Bombay.}
    \fact{Oct 2025}{Global Firepower Index 2025: USA 1, Russia 2, China 3, India 4; PowerIndex score 0.1184.}
    \fact{Nov 2025}{India Skills Report 2026: employability 56.35\%; women 54\%, men 51.5\%; 22--25 age group 75.7\%.}
    \fact{Nov--Dec 2025}{Rule of Law Index 2025: India rank 86/143, Denmark topped; Arton Passport Index 2025: UAE rank 1, India rank 67.}
    \fact{Jan 2026}{Henley Passport Index 2026: Singapore rank 1 with 192 destinations; India rank 80 with 55; Export Preparedness Index 2024 topped by Maharashtra.}
    \fact{Feb 2026}{Network Readiness Index 2025: India rank 45, USA topped; CPI 2025: India rank 91, Denmark rank 1; LEED list: India rank 2, China topped.}
    \fact{Mar 2026}{World Happiness Report 2026: Finland rank 1 for 9th year; India rank 116/147; Odisha rank 1 in Fiscal Health Index 2026.}
\end{itemize}

\subsection{National/International}
\begin{itemize}
    \fact{May 2025}{President invoked Article 143 on Governors'/President's assent powers under Articles 200 and 201 after Tamil Nadu Governor delay case.}
    \fact{Jun 2025}{Census 2027 announced in 2 phases with caste enumeration; reference date 1 March 2027, and 1 October 2026 for Ladakh/snow-bound J\&K, HP, Uttarakhand.}
    \fact{Jun 2025}{Chenab Rail Bridge inaugurated: world's highest railway arch bridge, 359 m above Chenab, 467 m arch span; USBRL length 272 km.}
    \fact{Jul 2025}{India-UK FTA signed on 24 July; 99\% of Indian exports to UK duty-free; whisky/gin duty from 150\% to 75\% initially and 40\% over 10 years.}
    \fact{Sep 2025}{C.P. Radhakrishnan elected 15th Vice President with 452 first-preference votes; Justice B. Sudershan Reddy got 300.}
    \fact{Nov 2025}{20th G20 Summit, Johannesburg, theme ``Solidarity, Equality, Sustainability''; 122-point Leaders' Declaration; US did not participate.}
    \fact{Dec 2025}{23rd India-Russia Annual Summit: Programme 2030 set USD 100 billion trade target; Brazil transferred BRICS Presidency to India for 2026.}
    \fact{Jan 2026}{India-UAE target: bilateral trade USD 200 billion by 2032; LoIs included Dholera SIR, space, defence and 0.5 MMTPA LNG for 10 years from 2028.}
    \fact{Feb 2026}{India hosted 2nd India-Arab Foreign Ministers Meeting in New Delhi with 22 Arab League countries; New Delhi Declaration adopted.}
    \fact{Apr 2026}{Amaravati declared sole capital of Andhra Pradesh by Gazette notification dated 6 April 2026; Women's Reservation Bill failed despite 298--230 vote, below 2/3 threshold of 360.}
\end{itemize}

\subsection{Environment/Health}
\begin{itemize}
    \fact{May 2025}{WHO adopted first global pandemic treaty at WHA, Geneva: 124 in favour, 0 against, 11 abstentions; focus equity, One Health and pathogen benefit sharing.}
    \fact{May 2025}{Operation Olivia protected 6.98 lakh Olive Ridley turtles at Rushikulya, Odisha; 16th Asiatic Lion Census showed 32\% rise since 2020.}
    \fact{Jun 2025}{``Ek Ped Maa Ke Naam 2.0'' target 10 crore trees by 30 September 2025; Kerala renamed Aralam as India's first butterfly sanctuary with 266+ species.}
    \fact{Sep 2025}{Cold Desert Biosphere Reserve became India's 13th UNESCO World Network biosphere reserve; Bihar added 2 Ramsar sites, taking India to 93.}
    \fact{Oct 2025}{Project Dolphin Phase II, Project Sloth Bear and Project Gharial launched; Palk Bay recognised as India's first dugong reserve by IUCN.}
    \fact{Nov 2025}{COP30 held in Belem, Brazil; Belem Package adopted with 29 decisions; Global Big Cats Summit 2026 announced for New Delhi.}
    \fact{Dec 2025}{Deepavali added to UNESCO Representative List at 20th ICH Committee session at Red Fort; Chhattisgarh's Kopra Reservoir became state's first Ramsar Site.}
    \fact{Jan 2026}{India non-fossil capacity reached 266.78 GW, solar 131.81 GW; Nutrient-Based Subsidy Rabi 2025--26 requirement Rs 37,952.29 crore.}
    \fact{Feb 2026}{WHO prequalified nOPV2 by Biological E; Supreme Court declared Menstrual Hygiene Management a Fundamental Right under Article 21.}
    \fact{Apr 2026}{Goldman Environmental Prize 2026 first all-women cohort; each winner USD 200,000; Solid Waste Management Rules 2026 mandated four-way segregation.}
\end{itemize}

\subsection{Science-Tech}
\begin{itemize}
    \fact{May 2025}{6th semiconductor unit approved: HCL-Foxconn JV near Jewar, Rs 3,700 crore, 20,000 wafers/month and 36 million units/month.}
    \fact{Jun 2025}{Shubhanshu Shukla piloted Axiom-4 to ISS on 25 June; 2nd Indian in space after Rakesh Sharma and first Indian to live/work on ISS.}
    \fact{Jul--Aug 2025}{BITS Pilani AI Plus Campus in Amaravati: Rs 1,000 crore; semiconductor projects Rs 4,600 crore in Odisha, Punjab and Andhra Pradesh; CG Power Sanand OSAT Rs 7,600 crore.}
    \fact{Sep 2025}{Vikram-3201, first indigenous 32-bit microprocessor for space use, unveiled; Semicon India 2025 showed 10 projects worth Rs 1.5 lakh crore.}
    \fact{Oct 2025}{Nobel 2025: Medicine for peripheral immune tolerance; Physics for macroscopic quantum tunnelling; Chemistry for metal-organic frameworks; Peace to Maria Corina Machado.}
    \fact{Nov--Dec 2025}{BIRSA 101 became first indigenous CRISPR gene therapy for sickle cell disease; DHRUV64, India's first indigenous 1.0 GHz 64-bit dual-core microprocessor, unveiled.}
    \fact{Jan 2026}{PSLV-C62/EOS-N1 launched on 12 January; anomaly in PS3; mission was PSLV's 64th flight and 105th launch from Sriharikota.}
    \fact{Feb--Mar 2026}{India AI Impact Summit at Bharat Mandapam: 100+ countries; M.A.N.A.V. framework; declaration endorsed by 92 countries/organisations and 2.5 lakh pledges set Guinness record.}
    \fact{Mar 2026}{IndiaAI compute capacity: 38,000+ GPUs provisioned, 20,000 more to be added; summit generated over USD 200 billion AI investment commitments.}
    \fact{Apr 2026}{NARIT AI launched by Gujarat Police; Gujarat HC barred AI in judgment drafting; Education Ministry introduced AI \& Computational Thinking for Classes 3--8.}
\end{itemize}

\subsection{Defence}
\begin{itemize}
    \fact{May/Aug 2025}{Operation Sindoor struck 9 terror sites linked to LeT/JeM/Hizbul; IAF hit 4 targets inside Pakistan and 5 in PoK; Navy supported from Indian soil.}
    \fact{May 2025}{BrahMos production facility inaugurated at UP Defence Industrial Corridor; target 80--100 BrahMos missiles annually.}
    \fact{Jun--Sep 2025}{Tiger Claw: first IAF Garud-US special forces exercise in UP; Yudh Abhyas 2025 held at Fort Wainwright, Alaska, 1--14 September.}
    \fact{Oct 2025}{India-US signed 10-year defence framework in Kuala Lumpur on 31 October; President Murmu became first Indian President to fly in Rafale, 30-minute sortie from Ambala.}
    \fact{Nov 2025}{US approved 100 FGM-148 Javelin ATGMs worth USD 45.7 million and 216 Excalibur rounds worth USD 47.1 million for India.}
    \fact{Dec 2025}{MoD signed Rs 7,995 crore, 5-year FMS support agreement with USA for 24 MH-60R Seahawk helicopters; QUAD FTX held at Guam, 8--12 December.}
    \fact{Jan 2026}{Army declared 2026 ``Year of Networking \& Data Centricity''; Radio Sangam launched at Keri village, Rajouri, on 88.8 FM.}
    \fact{Feb 2026}{MHA launched PRAHAAR, National Counter-Terrorism Policy based on seven pillars; Yantra India Limited got Miniratna Category-I status.}
    \fact{Apr 2026}{Navy inducted INS Aridaman, third Arihant-class SSBN; about 7,000-tonne displacement, K-15 and K-4 missiles, K-4 range up to 3,500 km.}
\end{itemize}

\subsection{Sports}
\begin{itemize}
    \fact{May--Jun 2025}{Neeraj Chopra crossed 90 m with 90.23 m at Doha Diamond League and won Paris Diamond League; RCB won maiden IPL 2025 title.}
    \fact{Jun--Jul 2025}{Carlos Alcaraz won French Open 2025; Jannik Sinner won first Wimbledon title; Iga Swiatek won first Wimbledon title.}
    \fact{Jul 2025}{Divya Deshmukh won FIDE Women's Chess World Cup in Batumi, defeating Koneru Humpy; became India's 4th woman Grandmaster.}
    \fact{Sep 2025}{India won 9th Asia Cup title by beating Pakistan by 5 wickets in Dubai; Tilak Varma Man of Match, Kuldeep Yadav MVP, Abhishek Sharma Player of Tournament.}
    \fact{Nov 2025}{ISSF World Championship Rifle/Pistol, Cairo: India 3rd with 13 medals, 3 gold/6 silver/4 bronze; China topped with 21.}
    \fact{Dec 2025}{ISSF World Cup Final, Doha: China topped with 9 medals, India 2nd with 6; Simranpreet Kaur and Suruchi Singh won golds.}
    \fact{Feb 2026}{WPL 2026: RCB beat Delhi Capitals by 6 wickets at Vadodara; Smriti Mandhana Player of Match, Orange Cap; Sophie Devine Purple Cap and MVP.}
    \fact{Feb 2026}{Australian Open 2026: Carlos Alcaraz men's singles, Elena Rybakina women's singles; Milano Cortina Winter Olympics held in Italy, 6--22 February.}
    \fact{Mar 2026}{ICC Men's T20 World Cup 2026 final, Ahmedabad: India beat New Zealand by 96 runs; India 255/5, NZ 159; Sanju Samson 89, Bumrah 4/15.}
    \fact{Apr 2026}{Laureus 2026, Madrid: Sportsman Carlos Alcaraz, Sportswoman Aryna Sabalenka, Team Paris Saint-Germain, Breakthrough Lando Norris.}
\end{itemize}

\subsection{Awards/Appointments}
\begin{itemize}
    \fact{May 2025}{Pulitzer 2025: Public Service to Kavitha Surana, Lizzie Presser, Cassandra Jaramillo and Stacy Kranitz; Reuters won Investigative Reporting for ``Fentanyl Express''.}
    \fact{Jun--Jul 2025}{PM Modi received Cyprus' Grand Cross of the Order of Makarios III, Ghana's Officer of the Order of the Star, Trinidad \& Tobago's Order of the Republic, Namibia's Welwitschia Mirabilis, and Brazil's Grand Collar.}
    \fact{Aug/Sep 2025}{Educate Girls became first Indian organisation to win Ramon Magsaysay Award 2025; founded by Safeena Husain; 2 million+ girls enrolled/retained.}
    \fact{Sep 2025}{71st National Film Awards: Dadasaheb Phalke to Mohanlal; Best Feature Film \textit{12th Fail}; Best Actors Shah Rukh Khan and Vikrant Massey; Best Actress Rani Mukerji.}
    \fact{Oct 2025}{70th Filmfare Awards: \textit{Laapataa Ladies} won 13 awards including Best Film and Best Director Kiran Rao; Justice Surya Kant named 53rd CJI.}
    \fact{Dec 2025}{IFFI Goa: Golden Peacock to \textit{Skin of Youth}; Satyajit Ray Lifetime Achievement Award to Rajinikanth; Japan Country of Focus.}
    \fact{Jan--Feb 2026}{Rakesh Aggarwal appointed NIA DG till August 2028; Golden Globes Best Film Drama \textit{Hamnet}; Grammy Song of the Year Billie Eilish's ``Wildflower''.}
    \fact{Mar 2026}{Oscars 2026: Best Picture \textit{One Battle After Another}; Best Actor Michael B. Jordan; Best Actress Jessie Buckley; Best Director Paul Thomas Anderson.}
    \fact{Mar 2026}{Sahitya Akademi Award 2025 English winner Navtej Sarna for \textit{Crimson Spring}; prize Rs 1,00,000, shawl and engraved copper plaque.}
    \fact{Apr 2026}{Chetak Screen Awards: Best Film \textit{Homebound}, Best Director Aditya Dhar, Best Actor Ranveer Singh, Best Actress Yami Gautam; Ashwini Bhide became first woman BMC Commissioner.}
\end{itemize}

\subsection{Summits/Important Days}
\begin{itemize}
    \fact{May 2025}{10th NITI Aayog Governing Council Meeting at Bharat Mandapam; theme ``Viksit Rajya for Viksit Bharat @2047''; 24 states and 7 UTs attended.}
    \fact{May--Jun 2025}{Important days: National Technology Day 11 May; International Nurses Day 12 May; World Environment Day 5 June; International Olympic Day 23 June.}
    \fact{Jul 2025}{World Population Day 11 July; World Youth Skills Day 15 July; World Chess Day 20 July; World Mangrove Day 26 July; World Hepatitis Day 28 July.}
    \fact{Sep 2025}{25th SCO Summit held in Tianjin; Tianjin Declaration condemned 22 April Pahalgam attack; World Food India 2025 held at Bharat Mandapam, 25--28 September.}
    \fact{Oct 2025}{Timor-Leste became ASEAN's 11th member at Kuala Lumpur Summit on 26 October; Janajatiya Gaurav Diwas falls on 15 November for Birsa Munda.}
    \fact{Nov 2025}{G20 Summit, Johannesburg, 22--23 November, theme ``Solidarity, Equality, Sustainability''; PM proposed six ideas including Global Health Response Team and Open Satellite Data Partnership.}
    \fact{Dec 2025}{20th UNESCO ICH Committee session at Red Fort, 8--13 December, first time in India; 67 new ICH elements added globally.}
    \fact{Jan 2026}{WEF Annual Meeting at Davos: theme ``A Spirit of Dialogue''; BRICS 2026 theme ``Building for Resilience, Innovation, Cooperation and Sustainability''.}
    \fact{Feb--Mar 2026}{India chaired first BRICS Sherpas' Meeting 2026 in New Delhi; Janaushadhi Saptah observed 1--7 March; National Science Day theme ``Women in Science: Catalysing Viksit Bharat''.}
    \fact{Apr 2026}{Samrat Samprati Museum inaugurated at Koba, Gandhinagar, on Mahavir Jayanti; nearly 2,000 Jain artefacts in 7 galleries.}
\end{itemize}

\subsection{Static GK traps}
\begin{itemize}
    \fact{May 2025}{India's official coastline revised from 7,516.6 km to 11,098.8 km; 11 coastal states and 2 UTs; measured by Survey of India and National Hydrographic Office.}
    \fact{May 2025}{Lakkundi monuments: Karnataka, Kalyana/Western Chalukya, 10th--12th century CE, Vesara style; key sites Kasi Visvesvara, Brahma Jinalaya, Musukina Bavi.}
    \fact{Jun 2025}{SCO members in source: India, China, Russia, Pakistan, Kazakhstan, Kyrgyzstan, Tajikistan, Uzbekistan, Iran, Belarus; Secretariat Beijing, RATS Tashkent.}
    \fact{Jul 2025}{Maratha Military Landscapes UNESCO entry: 12 forts; 11 in Maharashtra plus Gingee in Tamil Nadu; added at 47th WHC session in Paris.}
    \fact{Aug 2025}{Sawalkote Hydropower Project: 2,185 MW run-of-river on Chenab in J\&K, Rs 22,704 crore, target 7,000 million units annually.}
    \fact{Oct 2025}{\textit{Vande Mataram}: composed by Bankim Chandra Chattopadhyay, first sung by Rabindranath Tagore in 1896; 150 years noted.}
    \fact{Nov 2025}{G20 established 1999; first Leaders' Summit 2008; 19 countries plus EU and African Union; about 85\% global GDP, 75\% trade, two-thirds population.}
    \fact{Dec 2025}{UNESCO HQ Paris, established 1945; Khaled El-Enany is Director-General.}
    \fact{Feb 2026}{Arunachal Pradesh and Mizoram Statehood Day: 20 February; both attained statehood in 1987.}
    \fact{Apr 2026}{Nuclear triad: launch capability from land-based missiles, aircraft and sea-based submarines; supports second-strike capability.}
\end{itemize}
% END TOPIC MASTER

% BEGIN STATIC-CURRENT GK
\section{Static-Current GK: Source-Backed Trap Lists}

\subsection{Source Notes and Scope}
\begin{itemize}
    \item This section targets static facts that often ride with current affairs: full forms, headquarters, founders/heads, regulators, exchanges, banks, airports, international bodies, committees, and institutions.
    \item Source base used for this expansion: Adda247 Static GK hub, Adda247 Full Forms hub, Adda247 International Organizations and Headquarters, Adda247 Major Stock Exchanges in India, Adda247 Important Committees in India, PracticeMock Static GK PDF index, PracticeMock Indian Financial Regulators, PracticeMock Indian Banks and Headquarters, PracticeMock International Organisations and Headquarters, PracticeMock Famous Indian Airports, and PracticeMock Banking Abbreviations.
    \item Exam method: do not read these as essays. Convert every line into a match-the-following pair or a one-line MCQ: entity--HQ, abbreviation--full form, institution--year, bank--tagline, body--purpose.
\end{itemize}

\subsection{Banking, Finance and Regulator Traps}
\begin{itemize}
    \item RBI: established 1 April 1935 under RBI Act, 1934; original central office Kolkata, moved permanently to Mumbai in 1937; nationalised in 1949.
    \item NABARD: established 12 July 1982 on B. Sivaramman Committee recommendation; headquarters Mumbai; apex institution for agriculture and rural development finance.
    \item SEBI: established 12 April 1992 under SEBI Act, 1992; regulator of securities and commodity markets.
    \item SIDBI: set up 2 April 1990; headquarters Lucknow; principal financial institution for MSME promotion, financing and development.
    \item NHB: set up 9 July 1988 on C. Rangarajan group recommendation; headquarters New Delhi; apex housing finance institution.
    \item IRDAI: statutory insurance regulator; headquarters Hyderabad; 10-member body including chairperson, five full-time and four part-time members.
    \item DICGC: RBI subsidiary; established 15 July 1978; headquarters Mumbai; deposit insurance limit Rs 5,00,000 per depositor per bank.
    \item ECGC: export credit insurance support for Indian exporters; based in Mumbai; controlled by Ministry of Commerce.
    \item CCI: established 14 October 2003; enforces Competition Act, 2002; promotes fair competition and protects consumer interest.
\end{itemize}

\subsection{Banks, Headquarters and Taglines}
\begin{itemize}
    \item SBI -- Mumbai -- ``The Banker to Every Indian''.
    \item Punjab National Bank -- New Delhi -- ``The Name You Can Bank Upon''.
    \item Bank of Baroda -- Vadodara -- ``India's International Bank''.
    \item Canara Bank -- Bengaluru -- ``Together We Can''.
    \item Union Bank of India -- Mumbai -- ``Good People to Bank With''.
    \item Indian Bank -- Chennai -- ``Your Own Bank''.
    \item Bank of India -- Mumbai -- ``Relationship Beyond Banking''.
    \item Central Bank of India -- Mumbai -- ``Central to You Since 1911''.
    \item Indian Overseas Bank -- Chennai -- ``Good People to Grow With''.
    \item Punjab \& Sind Bank -- New Delhi -- ``Where Service is a Way of Life''.
    \item UCO Bank -- Kolkata -- ``Honors Your Trust''.
    \item Bank of Maharashtra -- Pune -- ``One Family One Bank''.
    \item Small finance banks to pair quickly: AU -- Jaipur; Utkarsh -- Varanasi; Equitas -- Chennai; ESAF -- Thrissur; Jana -- Bengaluru; Ujjivan -- Bengaluru; North East -- Guwahati; Suryoday -- Navi Mumbai.
    \item Payments banks to pair quickly: Airtel -- New Delhi; India Post Payments Bank -- New Delhi; Jio -- Mumbai; Fino -- Mumbai; Paytm -- Noida; NSDL -- Mumbai.
\end{itemize}

\subsection{Stock Exchanges and Market Infrastructure}
\begin{itemize}
    \item BSE: Bombay Stock Exchange; headquarters Mumbai; established 1875; Asia's oldest stock exchange.
    \item NSE: National Stock Exchange of India; headquarters Mumbai; established 1992; associated benchmark Nifty 50.
    \item Calcutta Stock Exchange: headquarters Kolkata; established 1908.
    \item India International Exchange (India INX): Gujarat; established 2017; linked with GIFT City/IFSC context.
    \item Metropolitan Stock Exchange of India: Mumbai; established 2008.
    \item Social Stock Exchange: platform for securities of social enterprises and impact-driven organisations; revise with SEBI/social-sector current affairs.
\end{itemize}

\subsection{International Organisations and Headquarters}
\begin{itemize}
    \item UN -- New York -- established 24 October 1945.
    \item WHO -- Geneva -- established 7 April 1948.
    \item IMF -- Washington DC -- established 27 December 1945.
    \item World Bank -- Washington DC -- Bretton Woods/July 1944 origin.
    \item WTO -- Geneva -- established 1 January 1995.
    \item UNESCO -- Paris -- established 16 November 1945.
    \item FAO -- Rome -- established 16 October 1945.
    \item ILO -- Geneva -- established 1919.
    \item WIPO -- Geneva -- established 14 July 1967.
    \item IAEA -- Vienna -- established 29 July 1957.
    \item OPEC -- Vienna -- established September 1960.
    \item NATO -- Brussels -- established 4 April 1949.
    \item ASEAN -- Jakarta -- established 8 August 1967.
    \item SAARC -- Kathmandu -- established 8 December 1985.
    \item ADB -- Mandaluyong, Philippines -- established 19 December 1966.
    \item AIIB -- Beijing -- established 16 January 2016.
    \item New Development Bank/BRICS Development Bank -- Shanghai -- established 15 July 2014.
    \item IOC -- Lausanne; FIFA -- Zurich; WEF -- Geneva; BIS -- Basel; WWF -- Gland; ICJ -- The Hague; ICC -- Dubai.
\end{itemize}

\subsection{Indian Airports: Static Pairs}
\begin{itemize}
    \item Indira Gandhi International Airport -- Delhi -- Delhi.
    \item Chhatrapati Shivaji Maharaj International Airport -- Mumbai -- Maharashtra.
    \item Kempegowda International Airport -- Bengaluru -- Karnataka.
    \item Chennai International Airport -- Chennai -- Tamil Nadu.
    \item Netaji Subhas Chandra Bose International Airport -- Kolkata -- West Bengal.
    \item Rajiv Gandhi International Airport -- Hyderabad -- Telangana.
\end{itemize}

\subsection{Full Forms and Banking Abbreviations}
\begin{itemize}
    \item NEFT -- National Electronic Funds Transfer.
    \item RTGS -- Real Time Gross Settlement.
    \item IMPS -- Immediate Payment Service.
    \item UPI -- Unified Payments Interface.
    \item NPCI -- National Payments Corporation of India.
    \item IFSC -- Indian Financial System Code.
    \item AEPS -- Aadhaar Enabled Payments System.
    \item VPA -- Virtual Payment Address.
    \item KYC -- Know Your Customer.
    \item NPA -- Non-Performing Asset.
    \item CRR -- Cash Reserve Ratio; SLR -- Statutory Liquidity Ratio; MSF -- Marginal Standing Facility; LAF -- Liquidity Adjustment Facility.
    \item NABARD -- National Bank for Agriculture and Rural Development; SIDBI -- Small Industries Development Bank of India; DICGC -- Deposit Insurance and Credit Guarantee Corporation.
    \item SEBI -- Securities and Exchange Board of India; IRDAI -- Insurance Regulatory and Development Authority of India; PFRDA -- Pension Fund Regulatory and Development Authority.
    \item GST -- Goods and Services Tax; GSTN -- Goods and Services Network; FDI -- Foreign Direct Investment; FPI -- Foreign Portfolio Investment; IPO -- Initial Public Offering.
    \item Tech-current full forms to revise from current-affairs/full-form pages whenever they appear in news: AI, API, LLM, GPT, QR, OTP, VPN, semiconductor/OSAT terms, and digital payment acronyms.
\end{itemize}

\subsection{Important Committees and Commissions}
\begin{itemize}
    \item Central Information Commission -- complaints under information law -- 2005.
    \item Central Vigilance Commission -- corruption prevention/investigation -- 1964.
    \item Election Commission of India -- free and fair elections -- 1950.
    \item Union Public Service Commission -- recruitment/personnel advice -- 1926.
    \item University Grants Commission -- university education standards -- 1956.
    \item National Human Rights Commission -- human-rights protection and promotion -- 1993.
    \item National Commission for Women -- women's rights protection and promotion -- 1992.
    \item Competition Commission of India -- fair competition and Competition Act enforcement -- 2003.
    \item Law Commission -- law review and reforms -- 1955.
    \item Finance Commission -- tax distribution between Centre and states -- 1951.
    \item Kothari Commission -- education policy -- 1964.
    \item Mandal Commission -- OBC reservation in jobs -- 1979.
    \item Sarkaria Commission -- Centre-state relations -- 1983.
    \item M.M. Punchhi Commission -- Centre-state relations -- 2007.
    \item Shah Commission -- Emergency excesses during 1975--77 -- 1977.
\end{itemize}

\subsection{How to Build Static-Current Flashcards}
\begin{itemize}
    \item For every company/institution appearing in monthly current affairs, add four fields: founder or head, headquarters, sector, and current event trigger.
    \item For every international summit or treaty, add three static fields: host city, organisation headquarters, and year of establishment.
    \item For every bank/RBI item, add three static fields: regulator, headquarters, and key abbreviation/full form.
    \item For every airport/airline/transport item, add route or location, headquarters/base city, and parent ministry/regulator if applicable.
    \item For every report/index, add publisher, headquarters of publisher, top rank, and India's rank or status.
\end{itemize}
% END STATIC-CURRENT GK

\section{Last-Week Drill}
\begin{enumerate}
    \item Write the RBI repo rate path and latest inflation/GDP forecasts from memory.
    \item List 20 reports/indices with publisher and India's rank or top performer.
    \item List 20 schemes with ministry/state and target group.
    \item List 20 awards with recipient and field.
    \item List 20 sports winners with venue and key player.
\end{enumerate}

\end{document}
